---
title: "ACP5-PondModel"
author: "Moritz Berg"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = "hide")
# Sys.setenv("R_REMOTES_NO_ERRORS_FROM_WARNINGS" = "true")
# remotes::install_github("GLEON/rLakeAnalyzer")
# remotes::install_github("aemon-j/GLM3r", ref = "v3.1.1")
# remotes::install_github("USGS-R/glmtools", ref = "ggplot_overhaul")
# remotes::install_github("aemon-j/FLakeR", ref = "inflow")
# remotes::install_github("aemon-j/GOTMr")
# remotes::install_github("aemon-j/gotmtools")
# remotes::install_github("aemon-j/SimstratR")
# remotes::install_github("aemon-j/MyLakeR")
# remotes::install_github("aemon-j/LakeEnsemblR")
library(rLakeAnalyzer)
library(GLM3r)
library(glmtools)
library(FLakeR)
library(GOTMr)
library(gotmtools)
library(SimstratR)
library(MyLakeR)
library(LakeEnsemblR)
library(readr)
library(lubridate)
library(tidyr)
library(dplyr)
library(readxl)
library(zoo)
library(data.table)
library(ggplot2)
```

# Task

2.  **Simulation of water temperature in scaled basins (original size of the mesocosms, x10, x100): What is the influence of basin size on water temperature and thermal stratification dynamics?**

Note: When answering the above-mentioned research questions, the focus should be on the methodological aspects of the respective model applications and the necessary analyses. A detailed answer to the questions in the context of other scientific work goes beyond the scope of this course.

Portfolio: Description of methods, results and a (short) discussion. The important thing is the complete documentation of the conducted model simulations.

## Data Preparation

The required format of data for the LakeEnsemblR model environment are csv tables containing the pond hypsograph, the Water Temperature Profile of the pond and the meteorological data. These files are prepared using the following code from the given datasets.

## Observed Data Preparation

```{r wtempprep}
# turn wide format of EERES Temperature to long format
EERES_Temperature = read_csv("Data/EERES_Temperature.csv")
EERES_T_long = pivot_longer(EERES_Temperature, cols=`30 cm`:`air`, names_to = "Depth_meter", values_to = "Water_Temperature_celsius") %>%
  filter(!Depth_meter=="air") %>%
  mutate(datetime = dmy_hm(time),
         Depth_meter =  gsub("[^0-9]", "", Depth_meter)) %>%
  mutate(Depth_meter = as.numeric(Depth_meter)/100) %>%
  select(-time) %>%
  select(datetime, Depth_meter, Water_Temperature_celsius) %>%
  filter(!is.na(Water_Temperature_celsius)) %>%
  mutate(datetime = format(floor_date(as_datetime(datetime), unit="10 minutes"),"%Y-%m-%d %H:%M:%S"))

EERES_T_dropNAcol = EERES_T_long %>%
  filter(Depth_meter!=0.15)
```

## Interpolation of missing data

To fill the observations the missing data at the depth of 15cm is interpolated from the temperature 5cm above and below.

```{r}
# Function to interpolate temperature for NA values at Depth_meter = 0.15
interpolate_na_temperature <- function(data) {
  # Filter for the specific datetime and depth
  temp_data <- data %>% filter(Depth_meter %in% c(0.1, 0.2) & !is.na(Water_Temperature_celsius))
  
  if (nrow(temp_data) == 2) {
    # Perform linear interpolation
    interpolated_temp <- approx(temp_data$Depth_meter, temp_data$Water_Temperature_celsius, xout = 0.15)$y
    return(interpolated_temp)
  } else {
    return(NA)  # Return NA if not enough data for interpolation
  }
}

# Apply interpolation only for NA values at Depth_meter = 0.15
EERES_T_long <- EERES_T_long %>%
  group_by(datetime) %>%
  mutate(Water_Temperature_celsius = ifelse(Depth_meter == 0.15 & is.na(Water_Temperature_celsius),
                                             interpolate_na_temperature(cur_data()),
                                             Water_Temperature_celsius)) %>%
  ungroup()

# write Water temperature to csv
write.table(EERES_T_long, file = "LakeEnsemblR_wtemp_profile_EERES.csv", sep = ",", dec = ".", row.names = FALSE, col.names = TRUE, quote = FALSE)
# write.table(EERES_T_dropNAcol, file = "LakeEnsemblR_wtemp_profile_EERES.csv", sep = ",", dec = ".", row.names = FALSE, col.names = TRUE, quote = FALSE)
```

## Meteo Data Preparation

Following code presents the preparation of the meteorological data. This is done by aggregation of the meteorological data to the temporal scale of the observed data (1min to 10min). This is done via averages, and sum of rainfall in mm per hour respectively. LakeEnsemblR requires an input of surface level barometric pressure in Pa, therefore the transformation from hPa was undertaken. The wind direction and speed at 10m above surface should be delivered. The exact location of the anemometer was not known, hence it was assumed that the given data reflects this requirement.

```{r meteoprep}
# Meteorological data
meteo_raw = read_xlsx("Data/20200101-20211231_Wetterstationsdaten_Andreas Lorke.xlsx", 
    skip = 1, na=c("-99999.9","-9999")) 
meteo = meteo_raw %>%
  as.data.frame() %>%
  mutate(across(where(is.numeric), ~ na_if(., -9999))) %>%      
  mutate(across(where(is.numeric), ~ na_if(., -99999.9))) %>%   # refactor NAs
  mutate(time=format(as_datetime(`~...1`),"%H:%M:%S"),
         date=format(as_datetime(`~...2`),"%Y-%m-%d")) %>%                      
  mutate(datetime=ymd(date) + hms(time))                        # datetime from date and time


# check for na values in all columns but the wind direction column WV
na_df <- meteo %>%
  select(SO,TE,FE,TP,WD,WR,RD,DR,RE) %>%
  filter(if_any(everything(), is.na))

# 10min meteo data

meteo_10m_t = meteo %>%
  # Melt the data to long format
  pivot_longer(cols = -c(datetime,`~...1`,`~...2`,time,date), names_to = "variable", values_to = "value") %>%
  # Create time groups
  mutate(time_group = floor_date(datetime, unit = "10 minutes")) %>%
  # Group by time_group and variable
  group_by(time_group, variable) %>%
  # Summarise the data to ensure a single value per group
  summarise(value = ifelse(variable == "RD", sum(value, na.rm = TRUE), mean(value, na.rm = TRUE)), 
            .groups = "drop") %>%
  # Gather back to wide format, ensuring unique values
  pivot_wider(names_from = variable, values_from = value, values_fn = list(value = mean))

# fill missing steps in time series meteo data
complete_time_series <- seq(from = min(meteo_10m_t$time_group, na.rm = TRUE), 
                  to = max(meteo_10m_t$time_group, na.rm = TRUE), 
                  by = "10 min")

meteo_complete <- data.frame(time_group=complete_time_series) %>%
  full_join(meteo_10m_t, by = "time_group")

meteo_filled <- meteo_complete %>%
#interpolate values
  mutate(across(-c(time_group, WV), ~ zoo::na.approx(.))) %>%
  select(time_group,SO,TE,FE,TP,WD,WR,RD,DR) %>%
#hPa to Pa
  mutate(DR = 100*DR) %>%
#mm/10min to mm/hr
  mutate(RD = 6*RD) %>%
#rename and restructure to LakeEnsemblR Standard
  rename(datetime=time_group,
  Shortwave_Radiation_Downwelling_wattPerMeterSquared=SO,
  Air_Temperature_celsius=TE,
  Relative_Humidity_percent=FE,
  Dewpoint_Temperature_celsius=TP,
  Ten_Meter_Elevation_Wind_Speed_meterPerSecond=WD,
  Ten_Meter_Elevation_Wind_Direction_degree=WR,
  Rainfall_millimeterPerHour=RD,
  Surface_Level_Barometric_Pressure_pascal=DR) %>%
  mutate(datetime = format(datetime, "%Y-%m-%d %H:%M:%S")) %>%
  select(datetime,Ten_Meter_Elevation_Wind_Speed_meterPerSecond,Air_Temperature_celsius,Relative_Humidity_percent,Shortwave_Radiation_Downwelling_wattPerMeterSquared,Surface_Level_Barometric_Pressure_pascal,Rainfall_millimeterPerHour)

# check for remaining NAs
meteo_filled %>%
  filter(if_any(everything(), is.na))

# check for missing time steps
all(meteo_filled$time_group == complete_time_series)

write.table(meteo_filled, file = "LakeEnsemblR_meteo_EERES.csv", sep = ",", dec = ".", row.names = FALSE, col.names = TRUE, quote = FALSE)
```

# Model

## Ensemble model run

Following Code runs the LER model. Output is already saved by moving ot folder output_FLake_GOTM.

```{r ensemblemodel}
getwd()
# setwd("../")
# Set config file & models
config_file <- "LakeEnsemblR.yaml"
# model <- c("FLake", "GLM", "GOTM", "Simstrat",  "MyLake")
model <- c("FLake","GOTM","Simstrat")

# Example run
# 1. Export settings - creates directories with all model setups and exports settings from the LER configuration file
export_config(config_file = config_file, model = model)

# 2. Run ensemble lake models
run_ensemble(config_file = config_file, model = model, parallel = T, return_list = T, ncores = 4)

#Check configs
check_master_config(config_file= config_file, model=model, exp_cnf = T)
get_template()

```

## Calibration

TODO: cali_ensemble doesn't work because of: `Fehler in seq.POSIXt(as.POSIXct(start, tz = tz), as.POSIXct(stop, tz = tz),  :    ungültige Zeichenkette für 'by'`

```{r}
calc_fit(ncdf = "output/ensemble_output.nc", model = model)

# run the calibration using the LHC method
cali_res <- cali_ensemble(config_file = config_file, num = 500, cmethod = "LHC",
                          parallel = TRUE, model = model)

# we can plot the results of the latin hypercube calibration for e.g. GOTM using plot_LHC
plot_LHC(config_file = config_file, model = "GOTM", res_files = cali_res$GOTM,
         qual_met = "nse")
# or we can plot it for all models using
plot_LHC(config_file = config_file, model = model, res_files = unlist(cali_res),
         qual_met = "nse", best = "high")

# load results of LHC in a data.frame
res_LHC <- load_LHC_results(config_file = config_file, model = model, res_files = unlist(cali_res))

# get the best parameter set for each model using rmse
best_p <- setNames(lapply(model, function(m)res_LHC[[m]][which.min(res_LHC[[m]]$rmse), ]), model)
print(best_p)
```

## Model runs of enlarged basin

TODO: Think about setting fix_wtlvl to false since this may be important for stratification dynamics. TODO: Set the initial water temperature on a scientifically based information (Use the known temperature and add lower temperatures at lower depths)

```{r scaledmodel}
# 10x depth ----
getwd()
config_file <- "LakeEnsemblR_10depth.yaml"
# model <- c("FLake", "GLM", "GOTM", "Simstrat",  "MyLake")
model <- c("FLake","GOTM","Simstrat")
export_config(config_file = config_file, model = model)
run_ensemble(config_file = config_file, model = model, parallel = T, return_list = T, ncores = 4)
file.rename(paste0(getwd(),"output"),
            paste0(getwd(),"output_10depth"))


# 10x area ----
config_file <- "LakeEnsemblR_10area.yaml"
# model <- c("FLake", "GLM", "GOTM", "Simstrat",  "MyLake")
model <- c("FLake","GOTM","Simstrat")
export_config(config_file = config_file, model = model)
run_ensemble(config_file = config_file, model = model, parallel = T, return_list = T, ncores = 4)
file.rename(paste0(getwd(),"output"),
            paste0(getwd(),"output_10area"))

# 10x area and 10x depth ----
config_file <- "LakeEnsemblR_10area10depth.yaml"
# model <- c("FLake", "GLM", "GOTM", "Simstrat",  "MyLake")
model <- c("FLake","GOTM","Simstrat")
export_config(config_file = config_file, model = model)
run_ensemble(config_file = config_file, model = model, parallel = T, return_list = T, ncores = 4)
file.rename(paste0(getwd(),"output"),
            paste0(getwd(),"output_10area10depth"))
```

## Single model runs

Following Code can be used to run single models and check for errors.

```{r singlemodelruns}
setwd("C:/Users/morit/Documents/ACP5_PondModel")
## Test FLake ----
# 1. Example - creates directories with all model setup
export_config(config_file = config_file, model = "FLake")

setwd("FLake")

FLakeR::run_flake(verbose = TRUE)
out <- read.table("output/output.dat", skip = 1, header = TRUE)
plot(out$time, out$Ts, type = "l")

setwd("../")

## Test GLM ----
# 1. Example - creates directories with all model setup
export_config(config_file = config_file, model = "GLM")

setwd("GLM")

GLM3r::run_glm(verbose = TRUE)
out <- glmtools::get_var("output/output.nc", "temp", reference = "surface")
plot(out$DateTime, out[, 2], type = "l")

setwd("../")

## Test GOTM ----
# 1. Example - creates directories with all model setup
export_config(config_file = config_file, model = "GOTM")

setwd("GOTM")

GOTMr::run_gotm(verbose = TRUE)
out <- gotmtools::get_vari("output/output.nc", "temp")
plot(out$time, out[, 2], type = "l")

setwd("../")

## Test Simstrat ----
# 1. Example - creates directories with all model setup
export_config(config_file = config_file, model = "Simstrat")

setwd("Simstrat")

SimstratR::run_simstrat(par_file = "simstrat.par", verbose = TRUE)
out <- read.table(file.path("output", "T_out.dat"), header = TRUE, 
                  sep = ",", check.names = FALSE)
plot(1:nrow(out), out[, ncol(out)], type = "l")

setwd("../")


## Test MyLake ----
# 1. Example - creates directories with all model setup
export_config(config_file = config_file, model = "MyLake")

# setwd("MyLake")
MyLakeR::run_mylake(sim_folder = ".", config_dat = "mylake.Rdata")
load("/MyLake/output/output.RData")
plot(res$tt, res$Tzt[1, ], type = "l")

setwd("../")

# END
```

# Analysis

Now we plot the heatmap of the ensemble run.

```{r ensembleheatmap}
## Plot model output using gotmtools/ggplot2
# Extract names of all the variables in netCDF
ncdf <- 'output/ensemble_output.nc'
vars <- gotmtools::list_vars(ncdf)
vars # Print variables

p1 <- plot_heatmap(ncdf)
# p1
# Change the theme and increase text size for saving
p1 <- p1 +
  theme_classic(base_size = 24) + 
  scale_colour_gradientn(limits = c(0, 21),
                         colours = rev(RColorBrewer::brewer.pal(11, "Spectral")))
# Save as a png file
p1
# ggsave('output/ensemble_heatmap.png', p1,  dpi = 300,width = 384,height = 280, units = 'mm')

```

TODO: not looked into this yet

```{r PlotEnsembleOUptput}
# Plot ensemble mean at 0.9m
model = c("FLake", "GLM", "GOTM", "Simstrat", "MyLake")
plot_ensemble(ncdf = ncdf, model = model, var = "temp", depth = 0.9)

# Load watertemp from netCDF file as a list
wtemp <- load_var(ncdf = ncdf, var = "temp", return = "list")
names(wtemp)

# Plot residual diagnostic plots
plots <- plot_resid(var_list = wtemp) # Plot residuals - returns a list of plots
names(plots) #

# Plot residuals vs. depth ("res_depth")
plots[[2]]

# Analyse the netCDF output
out <- analyse_ncdf(ncdf, model = model, spin_up = 0)
names(out)
str <- out[["strat"]]
str[str$year == 2010, ]
stats <- out[["stats"]]
stats
```
