---
title: "ACP5-PondModel"
author: "Moritz Berg"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = "hide")
# Sys.setenv("R_REMOTES_NO_ERRORS_FROM_WARNINGS" = "true")
# remotes::install_github("GLEON/rLakeAnalyzer")
# remotes::install_github("aemon-j/GLM3r", ref = "v3.1.1")
# remotes::install_github("USGS-R/glmtools", ref = "ggplot_overhaul")
# remotes::install_github("aemon-j/FLakeR", ref = "inflow")
# remotes::install_github("aemon-j/GOTMr")
# remotes::install_github("aemon-j/gotmtools")
# remotes::install_github("aemon-j/SimstratR")
# remotes::install_github("aemon-j/MyLakeR")
# remotes::install_github("aemon-j/LakeEnsemblR")
library(rLakeAnalyzer)
library(GLM3r)
library(glmtools)
library(FLakeR)
library(GOTMr)
library(gotmtools)
library(SimstratR)
library(MyLakeR)
library(LakeEnsemblR)
library(readr)
library(lubridate)
library(tidyr)
library(dplyr)
library(readxl)
library(zoo)
library(data.table)
```

# Task

2.  Simulation of water temperature in scaled basins (original size of the mesocosms, x10, x100): What is the influence of basin size on water temperature and thermal stratification dynamics?

Note: When answering the above-mentioned research questions, the focus should be on the methodological aspects of the respective model applications and the necessary analyses. A detailed answer to the questions in the context of other scientific work goes beyond the scope of this course.

Portfolio: Description of methods, results and a (short) discussion. The important thing is the complete documentation of the conducted model simulations.

## Data Preparation

The required format of data for the LakeEnsemblR model environment are csv tables containing the pond hypsograph, the Water Temperature Profile of the pond and the meteorological data. These files are prepared using the following code from the given datasets. LakeEnsemblR requires an input of surface level barometric pressure in Pa, therefore the transformation from hPa was undertaken. The wind direction and speed at 10m above surface should be delivered. The exact location of the anemometer was not known, hence it was assumed that the given data reflects this requirement. For ease of computational effort, the meteorological data is aggregated from minutely observations to hourly averages, and sum of rainfall in mm per hour respectively. Since a few observations across some hours were missing

```{r dataprep}
# turn wide format of EERES Temperature to long format
EERES_Temperature = read_csv("Data/EERES_Temperature.csv")
EERES_T_long = pivot_longer(EERES_Temperature, cols=`30 cm`:`air`, names_to = "Depth_meter", values_to = "Water_Temperature_celsius") %>%
  filter(!Depth_meter=="air") %>%
  mutate(datetime = dmy_hm(time),
         Depth_meter =  gsub("[^0-9]", "", Depth_meter)) %>%
  mutate(Depth_meter = as.numeric(Depth_meter)/100) %>%
  select(-time) %>%
  select(datetime, Depth_meter, Water_Temperature_celsius) %>%
filter(!is.na(Water_Temperature_celsius))

EERES_T_dropNAcol = EERES_T_long %>%
  filter(Depth_meter!=0.15)

# write Water temperature to csv
# write.csv(EERES_T_long,"LakeEnsemblR_wtemp_profile_EERES.csv",row.names=FALSE)
write.table(EERES_T_dropNAcol, file = "LakeEnsemblR_wtemp_profile_EERES.csv", sep = ",", dec = ".", row.names = FALSE, col.names = TRUE, quote = FALSE)

# Meteorological data
meteo_raw = read_xlsx("Data/20200101-20211231_Wetterstationsdaten_Andreas Lorke.xlsx", 
    skip = 1, na=c("-99999.9","-9999")) 
meteo = meteo_raw %>%
  as.data.frame() %>%
  mutate(across(where(is.numeric), ~ na_if(., -9999))) %>%      
  mutate(across(where(is.numeric), ~ na_if(., -99999.9))) %>%   # refactor NAs
  mutate(time=format(as_datetime(`~...1`),"%H:%M:%S"),
         date=format(as_datetime(`~...2`),"%Y-%m-%d")) %>%                      
  mutate(datetime=ymd(date) + hms(time))                        # datetime from date and time


# check for na values in all columns but the wind direction column WV
na_df <- meteo %>%
  select(SO,TE,FE,TP,WD,WR,RD,DR,RE) %>%
  filter(if_any(everything(), is.na))


meteo_10m_t = meteo %>%
  # Melt the data to long format
  pivot_longer(cols = -c(datetime,`~...1`,`~...2`,time,date), names_to = "variable", values_to = "value") %>%
  # Create time groups
  mutate(time_group = floor_date(datetime, unit = "10 minutes") + minutes(5)) %>%
  # Group by time_group and variable
  group_by(time_group, variable) %>%
  # Summarise the data to ensure a single value per group
  summarise(value = ifelse(variable == "RD", sum(value, na.rm = TRUE), mean(value, na.rm = TRUE)), 
            .groups = "drop") %>%
  # Gather back to wide format, ensuring unique values
  pivot_wider(names_from = variable, values_from = value, values_fn = list(value = mean))

# fill missing steps in time series meteo data
complete_time_series <- seq(from = min(meteo_10m_t$time_group, na.rm = TRUE), 
                  to = max(meteo_10m_t$time_group, na.rm = TRUE), 
                  by = "10 min")

meteo_complete <- data.frame(time_group=complete_time_series) %>%
  full_join(meteo_10m_t, by = "time_group")

meteo_filled <- meteo_complete %>%
#interpolate values
  mutate(across(-c(time_group, WV), ~ zoo::na.approx(.))) %>%
  select(time_group,SO,TE,FE,TP,WD,WR,RD,DR) %>%
#hPa to Pa
  mutate(DR = 100*DR) %>%
#mm/10min to mm/hr
  mutate(RD = 6*RD) %>%
#rename and restructure to LakeEnsemblR Standard
  rename(datetime=time_group,
  Shortwave_Radiation_Downwelling_wattPerMeterSquared=SO,
  Air_Temperature_celsius=TE,
  Relative_Humidity_percent=FE,
  Dewpoint_Temperature_celsius=TP,
  Ten_Meter_Elevation_Wind_Speed_meterPerSecond=WD,
  Ten_Meter_Elevation_Wind_Direction_degree=WR,
  Rainfall_millimeterPerHour=RD,
  Surface_Level_Barometric_Pressure_pascal=DR) %>%
  mutate(datetime = format(datetime, "%Y-%m-%d %H:%M:%S")) %>%
  select(datetime,Ten_Meter_Elevation_Wind_Speed_meterPerSecond,Air_Temperature_celsius,Relative_Humidity_percent,Shortwave_Radiation_Downwelling_wattPerMeterSquared,Surface_Level_Barometric_Pressure_pascal,Rainfall_millimeterPerHour)

# check for remaining NAs
meteo_filled %>%
  filter(if_any(everything(), is.na))

# check for missing time steps
all(meteo_filled$time_group == complete_time_series)

write.table(meteo_filled, file = "LakeEnsemblR_meteo_EERES.csv", sep = ",", dec = ".", row.names = FALSE, col.names = TRUE, quote = FALSE)
```

## Example Model

```{r example}

# Load LakeEnsemblR
library(LakeEnsemblR)

getwd()

# Set config file & models
config_file <- "LakeEnsemblR.yaml"
model <- c("FLake", "GLM", "GOTM", "Simstrat", "MyLake")

# Example run
# 1. Export settings - creates directories with all model setups and exports settings from the LER configuration file
export_config(config_file = config_file, model = model)

# 2. Run ensemble lake models
run_ensemble(config_file = config_file, model = model, parallel = T)

```



TODO

```{r multi-core run}
 # for(i in seq_len(num_cores)){
 #   core_job <- i
 #   rstudioapi::jobRunScript(path = "run_LER_on_core.R",
 #                            importEnv = TRUE,
 #                            name = paste0("job_", i))
 # }
```


# Analysis

```{r}

# Load libraries for post-processing
library(ggplot2)

## Plot model output using gotmtools/ggplot2
# Extract names of all the variables in netCDF
ncdf <- 'output/ensemble_output.nc'
vars <- gotmtools::list_vars(ncdf)
vars # Print variables

p1 <- plot_heatmap(ncdf)
p1
# Change the theme and increase text size for saving
p1 <- p1 +
  theme_classic(base_size = 24) + 
  scale_colour_gradientn(limits = c(0, 21),
                         colours = rev(RColorBrewer::brewer.pal(11, "Spectral")))
# Save as a png file
ggsave('output/ensemble_heatmap.png', p1,  dpi = 300,width = 384,height = 280, units = 'mm')

```

```{r calibration}
calc_fit(ncdf = "output/ensemble_output.nc", model = model)

# run the calibration using the LHC method
cali_res <- cali_ensemble(config_file = config_file, num = 500, cmethod = "LHC",
                          parallel = TRUE, model = model)

# we can plot the results of the latin hypercube calibration for e.g. GOTM using plot_LHC
plot_LHC(config_file = config_file, model = "GOTM", res_files = cali_res$GOTM,
         qual_met = "nse")
# or we can plot it for all models using
plot_LHC(config_file = config_file, model = model, res_files = unlist(cali_res),
         qual_met = "nse", best = "high")

# load results of LHC in a data.frame
res_LHC <- load_LHC_results(config_file = config_file, model = model, res_files = unlist(cali_res))

# get the best parameter set for each model using rmse
best_p <- setNames(lapply(model, function(m)res_LHC[[m]][which.min(res_LHC[[m]]$rmse), ]), model)
print(best_p)
```


```{r PlotEnsembleOUptput}
# Plot ensemble mean at 0.9m
model = c("FLake", "GLM", "GOTM", "Simstrat", "MyLake")
plot_ensemble(ncdf = ncdf, model = model, var = "temp", depth = 0.9)

# Load watertemp from netCDF file as a list
wtemp <- load_var(ncdf = ncdf, var = "temp", return = "list")
names(wtemp)

# Plot residual diagnostic plots
plots <- plot_resid(var_list = wtemp) # Plot residuals - returns a list of plots
names(plots) #

# Plot residuals vs. depth ("res_depth")
plots[[2]]

# Analyse the netCDF output
out <- analyse_ncdf(ncdf, model = model, spin_up = 0)
names(out)
str <- out[["strat"]]
str[str$year == 2010, ]
stats <- out[["stats"]]
stats
```

