---
title: "ACP5-PondModel"
author: "Moritz Berg"
date: "`r Sys.Date()`"
output: 
  html_document: 
    df_print: tibble
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = TRUE)
Sys.setenv("R_REMOTES_NO_ERRORS_FROM_WARNINGS" = "true")
# remotes::install_github("GLEON/rLakeAnalyzer")
# remotes::install_github("aemon-j/GLM3r", ref = "v3.1.1")
# remotes::install_github("USGS-R/glmtools", ref = "ggplot_overhaul")
# remotes::install_github("aemon-j/FLakeR", ref = "inflow")
# remotes::install_github("aemon-j/GOTMr")
# remotes::install_github("aemon-j/gotmtools")
# remotes::install_github("aemon-j/SimstratR")
# remotes::install_github("aemon-j/MyLakeR")
remotes::install_github("aemon-j/LakeEnsemblR")
# install.packages("C:/Users/morit/Documents/LakeEnsemblR", repos = NULL, type="source")
library(readr)
library(lubridate)
library(dplyr)
library(tidyr) 
library(readr)
library(readxl)
library(zoo)
library(data.table)
library(ggplot2)
library(gotmtools)
library(LakeEnsemblR)
```

# Task

**Validation of the models against observational data. Random and Systematic Discrepancies.**

Note: When answering the above-mentioned research questions, the focus should be on the methodological aspects of the respective model applications and the necessary analyses. A detailed answer to the questions in the context of other scientific work goes beyond the scope of this course.

Portfolio: Description of methods, results and a (short) discussion. The important thing is the complete documentation of the conducted model simulations.

## Data Preparation

The required format of data for the LakeEnsemblR model environment are csv tables containing the pond hypsograph, the Water Temperature Profile of the pond and the meteorological data. These files are prepared using the following code from the given datasets.

## Observed Data Preparation

The

```{r wtempprep}
# turn wide format of EERES Temperature to long format
EERES_Temperature = read_csv("Data/EERES_Temperature.csv")
EERES_T_long = pivot_longer(EERES_Temperature, cols=`30 cm`:`air`, names_to = "Depth_meter", values_to = "Water_Temperature_celsius") %>%
  filter(!Depth_meter=="air") %>%
  mutate(datetime = dmy_hm(time),
         Depth_meter =  gsub("[^0-9]", "", Depth_meter)) %>%
  mutate(Depth_meter = as.numeric(Depth_meter)/100) %>%
  select(-time) %>%
  select(datetime, Depth_meter, Water_Temperature_celsius) %>%
  filter(!is.na(Water_Temperature_celsius)) %>%
  mutate(datetime = format(floor_date(as_datetime(datetime), unit="10 minutes"),"%Y-%m-%d %H:%M:%S"))

# Optionally discard the Temperature at 15cm depth due to missing data
# EERES_T_dropNAcol = EERES_T_long %>%
#   filter(Depth_meter!=0.15)
```

## Interpolation of missing data

To fill the observations the missing data at the depth of 15cm is interpolated from the temperature 5cm above and below.

```{r}
# Function to interpolate temperature for NA values at Depth_meter = 0.15
interpolate_na_temperature <- function(data) {
  # Filter for the specific datetime and depth
  temp_data <- data %>% filter(Depth_meter %in% c(0.1, 0.2) & !is.na(Water_Temperature_celsius))
  
  if (nrow(temp_data) == 2) {
    # Perform linear interpolation
    interpolated_temp <- approx(temp_data$Depth_meter, temp_data$Water_Temperature_celsius, xout = 0.15)$y
    return(interpolated_temp)
  } else {
    return(NA)  # Return NA if not enough data for interpolation
  }
}

# Apply interpolation only for NA values at Depth_meter = 0.15
EERES_T_long <- EERES_T_long %>%
  group_by(datetime) %>%
  mutate(Water_Temperature_celsius = ifelse(Depth_meter == 0.15 & is.na(Water_Temperature_celsius),
                                             interpolate_na_temperature(cur_data()),
                                             Water_Temperature_celsius)) %>%
  ungroup()

# write Water temperature to csv
write.table(EERES_T_long, file = "LakeEnsemblR_wtemp_profile_EERES.csv", sep = ",", dec = ".", row.names = FALSE, col.names = TRUE, quote = FALSE)
# write.table(EERES_T_dropNAcol, file = "LakeEnsemblR_wtemp_profile_EERES.csv", sep = ",", dec = ".", row.names = FALSE, col.names = TRUE, quote = FALSE)
```

## Meteo Data Preparation

Following code presents the preparation of the meteorological data. This is done by aggregation of the meteorological data to the temporal scale of the observed data (1min to 10min). This is done via averages, and sum of rainfall in mm per hour respectively. LakeEnsemblR requires an input of surface level barometric pressure in Pa, therefore the transformation from hPa was undertaken. The wind direction and speed at 10m above surface should be delivered. The exact location of the anemometer was not known, hence it was assumed that the given data reflects this requirement.

```{r meteoprep}
# Meteorological data
meteo_raw = read_xlsx("Data/20200101-20211231_Wetterstationsdaten_Andreas Lorke.xlsx", 
    skip = 1, na=c("-99999.9","-9999")) 
meteo = meteo_raw %>%
  as.data.frame() %>%
  mutate(across(where(is.numeric), ~ na_if(., -9999))) %>%      
  mutate(across(where(is.numeric), ~ na_if(., -99999.9))) %>%   # refactor NAs
  mutate(time=format(as_datetime(`~...1`),"%H:%M:%S"),
         date=format(as_datetime(`~...2`),"%Y-%m-%d")) %>%                      
  mutate(datetime=ymd(date) + hms(time))                        # datetime from date and time


# check for na values in all columns but the wind direction column WV
na_df <- meteo %>%
  select(SO,TE,FE,TP,WD,WR,RD,DR,RE) %>%
  filter(if_any(everything(), is.na))

# 10min meteo data

meteo_10m_t = meteo %>%
  # Melt the data to long format
  pivot_longer(cols = -c(datetime,`~...1`,`~...2`,time,date), names_to = "variable", values_to = "value") %>%
  # Create time groups
  mutate(time_group = floor_date(datetime, unit = "10 minutes")) %>%
  # Group by time_group and variable
  group_by(time_group, variable) %>%
  # Summarise the data to ensure a single value per group
  summarise(value = ifelse(variable == "RD", sum(value, na.rm = TRUE), mean(value, na.rm = TRUE)), 
            .groups = "drop") %>%
  # Gather back to wide format, ensuring unique values
  pivot_wider(names_from = variable, values_from = value, values_fn = list(value = mean))

# fill missing steps in time series meteo data
complete_time_series <- seq(from = min(meteo_10m_t$time_group, na.rm = TRUE), 
                  to = max(meteo_10m_t$time_group, na.rm = TRUE), 
                  by = "10 min")

meteo_complete <- data.frame(time_group=complete_time_series) %>%
  full_join(meteo_10m_t, by = "time_group")

meteo_filled <- meteo_complete %>%
#interpolate values
  mutate(across(-c(time_group, WV), ~ zoo::na.approx(.))) %>%
  select(time_group,SO,TE,FE,TP,WD,WR,RD,DR) %>%
#hPa to Pa
  mutate(DR = 100*DR) %>%
#mm/10min to mm/hr
  mutate(RD = 6*RD) %>%
#rename and restructure to LakeEnsemblR Standard
  rename(datetime=time_group,
  Shortwave_Radiation_Downwelling_wattPerMeterSquared=SO,
  Air_Temperature_celsius=TE,
  Relative_Humidity_percent=FE,
  Dewpoint_Temperature_celsius=TP,
  Ten_Meter_Elevation_Wind_Speed_meterPerSecond=WD,
  Ten_Meter_Elevation_Wind_Direction_degree=WR,
  Rainfall_millimeterPerHour=RD,
  Surface_Level_Barometric_Pressure_pascal=DR) %>%
  mutate(datetime = format(datetime, "%Y-%m-%d %H:%M:%S")) %>%
  select(datetime,Ten_Meter_Elevation_Wind_Speed_meterPerSecond,Air_Temperature_celsius,Relative_Humidity_percent,Shortwave_Radiation_Downwelling_wattPerMeterSquared,Surface_Level_Barometric_Pressure_pascal,Rainfall_millimeterPerHour)

# check for remaining NAs
meteo_filled %>%
  filter(if_any(everything(), is.na))

# check for missing time steps
all(meteo_filled$time_group == complete_time_series)

write.table(meteo_filled, file = "LakeEnsemblR_meteo_EERES.csv", sep = ",", dec = ".", row.names = FALSE, col.names = TRUE, quote = FALSE)
```

```{r Nullmodeldata, results=hide, include=FALSE}
generate_random_meteo_data <- function(template, n) {
  sample_data <- data.frame(datetime = template$datetime)  # Keep datetime unchanged
  
  # Generate new data for each numeric column
  for (col in names(template)[-1]) {  # Exclude the datetime column
    if (is.numeric(template[[col]])) {
      # Get min and max values for the column
      min_val <- min(template[[col]], na.rm = TRUE)
      max_val <- max(template[[col]], na.rm = TRUE)
      
      # Calculate mean and standard deviation of the original column
      mean_val <- mean(template[[col]], na.rm = TRUE)
      sd_val <- sd(template[[col]], na.rm = TRUE)
      
      # Generate new values based on normal distribution
      new_values <- rnorm(n, mean = mean_val, sd = sd_val)
      
      # Ensure values are within the specified bounds
      new_values <- pmax(new_values, min_val)  # Set a lower bound
      new_values <- pmin(new_values, max_val)  # Set an upper bound
      
      sample_data[[col]] <- new_values
    }
  }
  
  return(sample_data)
}

# Generate a new sample dataframe based on the template
set.seed(123)  # For reproducibility
n_samples <- nrow(meteo_filled)  # Number of samples to generate
new_random_meteo_df <- generate_random_meteo_data(meteo_filled, n_samples)

# Function to generate new random data for Water_Temperature_celsius
generate_random_EERES_data <- function(template, n) {
  sample_data <- data.frame(datetime = template$datetime, Depth_meter = template$Depth_meter)  # Keep datetime and Depth unchanged
  
  # Get min and max values for Water_Temperature_celsius
  min_temp <- min(template$Water_Temperature_celsius, na.rm = TRUE)
  max_temp <- max(template$Water_Temperature_celsius, na.rm = TRUE)
  
  # Generate new temperature values
  mean_temp <- mean(template$Water_Temperature_celsius, na.rm = TRUE)
  sd_temp <- sd(template$Water_Temperature_celsius, na.rm = TRUE)
  
  # Generate new temperature values and ensure they are within the specified bounds
  new_temp_values <- rnorm(n, mean = mean_temp, sd = sd_temp)
  new_temp_values <- pmax(new_temp_values, min_temp)  # Ensure values are not below min_temp
  new_temp_values <- pmin(new_temp_values, max_temp)  # Ensure values are not above max_temp
  
  sample_data$Water_Temperature_celsius <- new_temp_values
  
  return(sample_data)
}


# Generate a new sample dataframe based on the template
set.seed(123)  # For reproducibility
n_samples <- nrow(EERES_T_long)  # Number of samples to generate
new_random_EERES_df <- generate_random_EERES_data(EERES_T_long, n_samples)

write.table(new_random_meteo_df, file = "LakeEnsemblR_meteo.csv", sep = ",", dec = ".", row.names = FALSE, col.names = TRUE, quote = FALSE)

write.table(new_random_EERES_df, file = "LakeEnsemblR_wtemp_profile.csv", sep = ",", dec = ".", row.names = FALSE, col.names = TRUE, quote = FALSE)
```

```{r Nullmodel, results=hide, include=FALSE}
getwd()
# setwd("../")
# Set config file & models
config_file <- "LakeEnsemblR.yaml"
# model <- c("FLake", "GLM", "GOTM", "Simstrat",  "MyLake")
model <- c("FLake","GOTM","Simstrat")

# Example run
# 1. Export settings - creates directories with all model setups and exports settings from the LER configuration file
export_config(config_file = config_file, model = model)

# 2. Run ensemble lake models
run_ensemble(config_file = config_file, model = model, parallel = T, return_list = T, ncores = 4)

ncdf <- 'output/ensemble_output.nc'
vars <- gotmtools::list_vars(ncdf)
vars # Print variables

p1 <- plot_heatmap(ncdf)
p1 <- p1 +
  theme_classic(base_size = 24) + 
  scale_colour_gradientn(limits = c(0, 21),
                         colours = rev(RColorBrewer::brewer.pal(11, "Spectral")))
p1

```

# Model

## Ensemble model run

Following Code runs the LER model. There are issues using GLM and MyLake in shallow ponds, therefore all following simulations are done using exclusively FLake, GOTM or Simstrat. We found that the model FLake calculates a mean depth of the lake from the volume (calculated from truncated square pyramids between the layers) and the surface area. This however seems to reduce the depth of shallow ponds below the depths of observed layers. Therefore we manually set the depth of the pond for FLake using the parameter `depth_w_lk`.

```{r ensemblemodel}
getwd()
# setwd("../")
# Set config file & models
config_file <- "LakeEnsemblR_EERES.yaml"
# model <- c("FLake", "GLM", "GOTM", "Simstrat",  "MyLake")
model <- c("FLake","GOTM","Simstrat")

# Example run
# 1. Export settings - creates directories with all model setups and exports settings from the LER configuration file
export_config(config_file = config_file, model = model)

# 2. Run ensemble lake models
run_ensemble(config_file = config_file, model = model, parallel = T, return_list = T, ncores = 3)

#Check configs
# check_master_config(config_file= config_file, model=model, exp_cnf = T)
```

## Calibration

The calibration is done using the function cali_ensemble. Due to errors by using this function on all models simultaneously, the results will be generated for each model solely. For each model, a calibration using a set of 100 parameter combinations is used. A Latin Hypercube Calibration (LHC) method is used in the calibration. With the function `plot_LHC`, the results can be shown graphically.

Depending on some arguments given to cali_ensemble (we guess `parallel`), the output of the function will either be the results or the names of the .csv files which store the parameter sets and results. Due to this inconsistency in the code, we manually select the generated results from the folder "cali".

```{r}
# getting the model fit of the last ensemble run
calc_fit(ncdf = "output/ensemble_output.nc", model = model)

config_file = "LakeEnsemblR_EERES_bestparams.yaml"

# # run the calibration using the LHC method
# cali_res_Simstrat <- cali_ensemble(config_file = config_file, num = 500, cmethod = "LHC",
#                           parallel = T, model = "SimStrat", ncores = 3)
# cali_res_GOTM <- cali_ensemble(config_file = config_file, num = 30, cmethod = "LHC",
#                           parallel = T, model = "GOTM", ncores = 3)
cali_res_FLake <- cali_ensemble(config_file = config_file, num = 30, cmethod = "LHC",
                          parallel = T, model = "FLake", ncores = 3)

# we manually set the file names with the results
res_file_GOTM <- paste0(getwd(),
                            c("/cali/params_GOTM_LHC_n500.csv",
                              "/cali/GOTM_LHC_n500.csv"))
res_file_FLake <- paste0(getwd(),
                            c("/cali/params_FLake_LHC_n100.csv",
                              "/cali/FLake_LHC_n100.csv"))
res_file_Simstrat <- paste0(getwd(),
                            c("/cali/params_Simstrat_LHC_n500.csv",
                              "/cali/Simstrat_LHC_n500.csv"))
# we can plot the results of the latin hypercube calibrations using plot_LHC
plot_LHC(config_file = config_file, model = "GOTM", res_files = res_file_GOTM,
         qual_met = "nse", best = "high")
# or we can plot it for all models using
plot_LHC(config_file = config_file, model = "FLake", res_files = res_file_FLake,
         qual_met = "nse", best = "high")
plot_LHC(config_file = config_file, model = "SimStrat", res_files = res_file_Simstrat,
         qual_met = "nse", best="high")
```

The results of the calibration can be extracted using `load_LHC_results`. Subsequently, the set of best parameters using the goodness of model fit measure Root Mean Squared Error (RMSE), Nash-Sutcliffe Efficiency (NSE) and mean absolute error (MAE) is printed.

```{r bestsets}
# load results of LHC in a data.frame
model = c("FLake","GOTM","Simstrat")
cali_res = list(res_file_FLake,res_file_GOTM,res_file_Simstrat)
res_LHC <- load_LHC_results(config_file = config_file, model = model, res_files = unlist(cali_res))

# Use res_LHC to investigate the results of the calibration

# get the best parameter set for each model using rmse
best_p_rmse <- setNames(lapply(model, function(m)res_LHC[[m]][which.min(res_LHC[[m]]$rmse), ]), model)
best_p_nse <- setNames(lapply(model, function(m)res_LHC[[m]][which.max(res_LHC[[m]]$nse), ]), model)
best_p_mae <- setNames(lapply(model, function(m)res_LHC[[m]][which.min(res_LHC[[m]]$mae), ]), model)
print("Set of parameters with lowest RMSE:")
print(best_p_rmse)
print("Set of parameters with highest NSE:")
print(best_p_rmse)
print("Set of parameters with lowest MAE:")
print(best_p_rmse)
```

We see that the best fitting parameter is independent of the quality metric used (RMSE, NSE or MAE). Subsequently, we will use this set of parameters in the following model. Since the parameters for wind speed (wind_speed) and shortwave radiation scaling (swr) are shared across all the models, the mean value of these from the best sets of parameters is calculated in the following code. The model specific parameters are printed below.

```{r bestparams}
# Initialize a list to store the mean values for each model
mean_values <- list()

# Initialize vectors to store wind_speed and swr values across all models
all_wind_speed <- c()
all_swr <- c()
all_kW <- c()

# Function to calculate mean values for a given model
calculate_means <- function(model_name) {
  # Extract dataframes for the model from each list
  df_rmse <- best_p_rmse[[model_name]]
  df_nse <- best_p_nse[[model_name]]
  df_mae <- best_p_mae[[model_name]]

  # Combine the dataframes
  combined_df <- rbind(df_rmse, df_nse, df_mae)

  # Calculate mean for the specified parameters
  mean_wind_speed <- mean(combined_df$wind_speed, na.rm = TRUE)
  mean_swr <- mean(combined_df$swr, na.rm = TRUE)
  mean_Kw <- mean(combined_df$Kw, na.rm = TRUE)

  # Check if the parameter exists in the combined dataframe before calculating mean
  mean_a_seiche <- ifelse("a_seiche" %in% colnames(combined_df),
                           mean(combined_df$a_seiche, na.rm = TRUE), NA)
  mean_turb_param_k_min <- ifelse("turb_param.k_min" %in% colnames(combined_df),
                                   mean(combined_df$turb_param.k_min, na.rm = TRUE), NA)
  mean_c_relax_C <- ifelse("c_relax_C" %in% colnames(combined_df),
                           mean(combined_df$c_relax_C, na.rm = TRUE), NA)

  # Append wind_speed and swr values to the global vectors
  all_wind_speed <<- c(all_wind_speed, combined_df$wind_speed)
  all_swr <<- c(all_swr, combined_df$swr)
  all_Kw <<- c(all_kW, combined_df$Kw)

  # Return the means as a named vector
  return(c(mean_wind_speed, mean_swr, mean_Kw, mean_a_seiche, mean_turb_param_k_min, mean_c_relax_C))
}

# Calculate means for each model
mean_values$FLake <- calculate_means("FLake")
mean_values$GOTM <- calculate_means("GOTM")
mean_values$Simstrat <- calculate_means("Simstrat")

# Convert the list to a dataframe for better visualization
mean_values_df <- do.call(rbind, mean_values)
colnames(mean_values_df) <- c("mean_wind_speed", "mean_swr", "mean_Kw", "mean_a_seiche", "mean_turb_param.k_min", "mean_c_relax_C")

# Calculate overall mean for wind_speed and swr across all models
overall_mean_wind_speed <- mean(all_wind_speed, na.rm = TRUE)
overall_mean_swr <- mean(all_swr, na.rm = TRUE)
overall_mean_Kw <- mean(all_Kw, na.rm=TRUE)

# Print the mean values for each model
cat("Parameter combination with best fit: \n")
print(mean_values_df)

# Print the overall mean values
cat("Overall mean wind_speed across all models:", overall_mean_wind_speed, "\n")
cat("Overall mean swr across all models:", overall_mean_swr, "\n")
cat("Overall mean Kw across all models:", overall_mean_Kw)
```

These parameters were implemented as initial parameters in a new configuration file (LakeEnsemblR_EERES_bestparams.yaml).

## Extending the calibration range

Now we expand the range of possible values since some of the best parameter sets have values close to the upper or lower bounds.

```{r, results='hide'}
config_file = "LakeEnsemblR_EERES_bestparams.yaml"
model = c("FLake","GOTM","Simstrat")
cali_res_GOTM_Simstrat_expanded <- cali_ensemble(config_file = config_file, num = 500, 
                                                 cmethod = "LHC", parallel = FALSE, 
                                                 model = c("GOTM","Simstrat"), 
                                                 ncores = 2)
cali_res_FLake_expanded <- cali_ensemble(config_file = config_file, num = 1,
                                         cmethod = "LHC", parallel = FALSE,
                                         model = "FLake")

plot_Simstrat = plot_LHC(config_file = config_file, 
         model = "Simstrat", 
         res_files = cali_res_GOTM_Simstrat_expanded$Simstrat,
         qual_met = "nse", best = "high")
plot_GOTM = plot_LHC(config_file = config_file, 
         model = "GOTM", 
         res_files = cali_res_GOTM_Simstrat_expanded$GOTM,
         qual_met = "nse", best = "high")
plot_FLake = plot_LHC(config_file = config_file, 
                      res_files = cali_res_FLake_expanded$FLake, 
                      model="FLake",
                      qual_met = "nse", best="high")

cali_res_all = list(cali_res_FLake,cali_res_GOTM,cali_res_Simstrat)
res_LHC_all <- load_LHC_results(config_file = config_file, model = model, res_files = unlist(cali_res_all))

# Use res_LHC to investigate the results of the calibration

# get the best parameter set for each model using rmse
best_p_rmse <- setNames(lapply(model, function(m)res_LHC_all[[m]][which.min(res_LHC_all[[m]]$rmse), ]), model)
best_p_nse <- setNames(lapply(model, function(m)res_LHC_all[[m]][which.max(res_LHC_all[[m]]$nse), ]), model)
best_p_mae <- setNames(lapply(model, function(m)res_LHC_all[[m]][which.min(res_LHC_all[[m]]$mae), ]), model)
```

## Model Run with best parameters

Followingly, the ensemble is run with the calibrated parameters. TODO: all.

```{r bestparamrun}
config_file = "LakeEnsemblR_EERES_calibrated.yaml"
model = c("FLake","GOTM","Simstrat")
export_config(config_file = config_file, model = model)
run_ensemble(config_file = config_file, model = model, parallel=T)
```

# Analysis

Now we plot the heatmap of the ensemble run.

```{r ensembleheatmap}
## Plot model output using gotmtools/ggplot2
# Extract names of all the variables in netCDF
ncdf <- 'output/ensemble_output.nc'
vars <- gotmtools::list_vars(ncdf)
vars # Print variables

p1 <- plot_heatmap(ncdf)
# p1
# Change the theme and increase text size for saving
p1 <- p1 +
  
  theme_classic(base_size = 24) + 
  scale_colour_gradientn(limits = c(0, 21),
                         colours = rev(RColorBrewer::brewer.pal(11, "Spectral")))
# Save as a png file
p1
# ggsave('output/ensemble_heatmap.png', p1,  dpi = 300,width = 384,height = 280, units = 'mm')

```

<<<<<<< HEAD
TODO: not looked into this yet

```{r PlotEnsembleOUptput}
# Plot ensemble mean at 0.9m
model = c("FLake", "GOTM", "Simstrat")
plot_ensemble(ncdf = ncdf, model = model, var = "temp", depth = 0.25)

# Load watertemp from netCDF file as a list
wtemp <- load_var(ncdf = ncdf, var = "temp", return = "list")
names(wtemp)

# Plot residual diagnostic plots
plots <- plot_resid(var_list = wtemp) # Plot residuals - returns a list of plots
names(plots) #

# Plot residuals vs. depth ("res_depth")
plots[[2]]

# Analyse the netCDF output
out <- analyse_ncdf(ncdf, model = model, spin_up = 0)
names(out)
str <- out[["strat"]]
str[str$year == 2010, ]
stats <- out[["stats"]]
stats
```

# Appendix
=======
```{r model results}
## Set up
ncdf <- "output_FLake_GOTM_Simstrat/ensemble_output.nc"
model = c("FLake", "GOTM", "Simstrat")

#### Results ####
# Take a look at the model fits to the observed data
error_metrics = calc_fit(ncdf = ncdf,
         model = model,
         var = "temp")



##### Model fits #####

##### Heatmap #####
p1 <- plot_heatmap(ncdf)

# Change the theme and increase text size for saving
p1 <- p1 +
  theme_classic(base_size = 24) + 
  scale_colour_gradientn(limits = c(0, 21),
                         colours = rev(RColorBrewer::brewer.pal(11, "Spectral")))

p1

##### Built-in: Models vs. Oberservations, Residuals, Temperature range boxplots (per depth) #####

# Plot ensemble mean at 0m
plot_ensemble(ncdf = ncdf, model = model, var = "temp", depth = 0, residuals = TRUE, boxwhisker = TRUE)

# Plot ensemble mean at 0.1m
plot_ensemble(ncdf = ncdf, model = model, var = "temp", depth = 0.1, residuals = TRUE, boxwhisker = TRUE)

# Plot ensemble mean at 0.15m
plot_ensemble(ncdf = ncdf, model = model, var = "temp", depth = 0.15, residuals = TRUE, boxwhisker = TRUE)

# Plot ensemble mean at 0.2m
plot_ensemble(ncdf = ncdf, model = model, var = "temp", depth = 0.2, residuals = TRUE, boxwhisker = TRUE)

# Plot ensemble mean at 0.25m
plot_ensemble(ncdf = ncdf, model = model, var = "temp", depth = 0.25, residuals = TRUE, boxwhisker = TRUE)

# Plot ensemble mean at 0.3m
# plot_ensemble(ncdf = ncdf, model = model, var = "temp", depth = 0.3, residuals = TRUE, boxwhisker = TRUE)



##### ggplot2: Water temperature predictions of models vs. observed temperatures (at depths in m: 0, 0.05, 0.1, 0.15, 0.2, 0.25) #####

# Load post-processed output data into your workspace
analyse_df <- analyse_ncdf(ncdf = ncdf, spin_up = NULL, drho = 0.1, model = model)


# Example plot the summer stratification period
strat_df <- analyse_df$strat
out_df <- analyse_df$out_df
obs <- analyse_df$obs_temp

obs$model = "Observed"
#obs$depths = obs$depths - 0.05
out_df = rbind(out_df, obs)


# Convert date column to POSIXct for proper time handling
out_df <- out_df %>%
  mutate(date = as.POSIXct(date, format = "%Y-%m-%d %H:%M:%S"),
         depth_label = paste0(depths, " m"))  # Create depth label

# Create the plot with labeled facets
ggplot(out_df, aes(x = date, y = temp, color = model)) +
  geom_line() +
  facet_wrap(~depth_label, scales = "free_y") +  # Label depth with "m"
  labs(title = "Water Temperature Predictions by Model",
       x = "Date", y = "Temperature (°C)", color = "Model") +
  theme_minimal() +
  theme(legend.position = "bottom",
        strip.text = element_text(size = 12))  # Make facet labels clearer




```
>>>>>>> 57b123ab197072dcc6eb57472d486dd1525279fa

## Single model runs

Following Code can be used to run single models and check for errors.

```{r singlemodelruns}
setwd("C:/Users/morit/Documents/ACP5_PondModel")
## Test FLake ----
# 1. Example - creates directories with all model setup
export_config(config_file = config_file, model = "FLake")

setwd("FLake")

FLakeR::run_flake(verbose = TRUE)
out <- read.table("output/output.dat", skip = 1, header = TRUE)
plot(out$time, out$Ts, type = "l")

setwd("../")

## Test GLM ----
# 1. Example - creates directories with all model setup
export_config(config_file = config_file, model = "GLM")

setwd("GLM")

GLM3r::run_glm(verbose = TRUE)
out <- glmtools::get_var("output/output.nc", "temp", reference = "surface")
plot(out$DateTime, out[, 2], type = "l")

setwd("../")

## Test GOTM ----
# 1. Example - creates directories with all model setup
export_config(config_file = config_file, model = "GOTM")

setwd("GOTM")

GOTMr::run_gotm(verbose = TRUE)
out <- gotmtools::get_vari("output/output.nc", "temp")
plot(out$time, out[, 2], type = "l")

setwd("../")

## Test Simstrat ----
# 1. Example - creates directories with all model setup
export_config(config_file = config_file, model = "Simstrat")

setwd("Simstrat")

SimstratR::run_simstrat(par_file = "simstrat.par", verbose = TRUE)
out <- read.table(file.path("output", "T_out.dat"), header = TRUE, 
                  sep = ",", check.names = FALSE)
plot(1:nrow(out), out[, ncol(out)], type = "l")

setwd("../")


## Test MyLake ----
# 1. Example - creates directories with all model setup
export_config(config_file = config_file, model = "MyLake")

# setwd("MyLake")
MyLakeR::run_mylake(sim_folder = ".", config_dat = "mylake.Rdata")
load("/MyLake/output/output.RData")
plot(res$tt, res$Tzt[1, ], type = "l")

setwd("../")

# END
```

## Model runs of enlarged basin

TODO: Think about setting fix_wtlvl to false since this may be important for stratification dynamics. TODO: Set the initial water temperature on a scientifically based information (Use the known temperature and add lower temperatures at lower depths)

```{r scaledmodel}

# 10x depth ----

getwd()

config_file <- "LakeEnsemblR_10depth.yaml"

# model <- c("FLake", "GLM", "GOTM", "Simstrat",  "MyLake")

model <- c("FLake","GOTM","Simstrat")

export_config(config_file = config_file, model = model)

run_ensemble(config_file = config_file, model = model, parallel = T, return_list = T, ncores = 4)

file.rename(paste0(getwd(),"output"),

            paste0(getwd(),"output_10depth"))

# 10x area ----

config_file <- "LakeEnsemblR_10area.yaml"

# model <- c("FLake", "GLM", "GOTM", "Simstrat",  "MyLake")

model <- c("FLake","GOTM","Simstrat")

export_config(config_file = config_file, model = model)

run_ensemble(config_file = config_file, model = model, parallel = T, return_list = T, ncores = 4)

file.rename(paste0(getwd(),"output"),

            paste0(getwd(),"output_10area"))

# 10x area and 10x depth ----

config_file <- "LakeEnsemblR_10area10depth.yaml"

# model <- c("FLake", "GLM", "GOTM", "Simstrat",  "MyLake")

model <- c("FLake","GOTM","Simstrat")

export_config(config_file = config_file, model = model)

run_ensemble(config_file = config_file, model = model, parallel = T, return_list = T, ncores = 4)

file.rename(paste0(getwd(),"output"),

            paste0(getwd(),"output_10area10depth"))

```
