---
title: "ACP5-PondModel"
author: "Moritz Berg"
date: "`r Sys.Date()`"
output: 
  html_document: 
    df_print: tibble
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = TRUE)
Sys.setenv("R_REMOTES_NO_ERRORS_FROM_WARNINGS" = "true")
# remotes::install_github("GLEON/rLakeAnalyzer")
# remotes::install_github("aemon-j/GLM3r", ref = "v3.1.1")
# remotes::install_github("USGS-R/glmtools", ref = "ggplot_overhaul")
# remotes::install_github("aemon-j/FLakeR", ref = "inflow")
# remotes::install_github("aemon-j/GOTMr")
# remotes::install_github("aemon-j/gotmtools")
# remotes::install_github("aemon-j/SimstratR")
# remotes::install_github("aemon-j/MyLakeR")
remotes::install_github("aemon-j/LakeEnsemblR")
# install.packages("C:/Users/morit/Documents/LakeEnsemblR", repos = NULL, type="source")
library(readr)
library(lubridate)
library(dplyr)
library(tidyr) 
library(readr)
library(readxl)
library(zoo)
library(data.table)
library(ggplot2)
library(gotmtools)
library(LakeEnsemblR)
library(patchwork)
library(randomForest)

```

# Task

**Validation of the models against observational data. Random and Systematic Discrepancies.**

Note: When answering the above-mentioned research questions, the focus should be on the methodological aspects of the respective model applications and the necessary analyses. A detailed answer to the questions in the context of other scientific work goes beyond the scope of this course.

Portfolio: Description of methods, results and a (short) discussion. The important thing is the complete documentation of the conducted model simulations.

## Data Preparation

The required format of data for the LakeEnsemblR model environment are csv tables containing the pond hypsograph, the Water Temperature Profile of the pond and the meteorological data. These files are prepared using the following code from the given datasets.

## Observed Data Preparation

The

```{r wtempprep}
# turn wide format of EERES Temperature to long format
EERES_Temperature = read_csv("Data/EERES_Temperature.csv")
EERES_T_long = pivot_longer(EERES_Temperature, cols=`30 cm`:`air`, names_to = "Depth_meter", values_to = "Water_Temperature_celsius") %>%
  filter(!Depth_meter=="air") %>%
  mutate(datetime = dmy_hm(time),
         Depth_meter =  gsub("[^0-9]", "", Depth_meter)) %>%
  mutate(Depth_meter = as.numeric(Depth_meter)/100) %>%
  select(-time) %>%
  select(datetime, Depth_meter, Water_Temperature_celsius) %>%
  filter(!is.na(Water_Temperature_celsius)) %>%
  mutate(datetime = format(floor_date(as_datetime(datetime), unit="10 minutes"),"%Y-%m-%d %H:%M:%S"))

# Optionally discard the Temperature at 15cm depth due to missing data
# EERES_T_dropNAcol = EERES_T_long %>%
#   filter(Depth_meter!=0.15)
```

## Interpolation of missing data

To fill the observations the missing data at the depth of 15cm is interpolated from the temperature 5cm above and below.

```{r}
# Function to interpolate temperature for NA values at Depth_meter = 0.15
interpolate_na_temperature <- function(data) {
  # Filter for the specific datetime and depth
  temp_data <- data %>% filter(Depth_meter %in% c(0.1, 0.2) & !is.na(Water_Temperature_celsius))
  
  if (nrow(temp_data) == 2) {
    # Perform linear interpolation
    interpolated_temp <- approx(temp_data$Depth_meter, temp_data$Water_Temperature_celsius, xout = 0.15)$y
    return(interpolated_temp)
  } else {
    return(NA)  # Return NA if not enough data for interpolation
  }
}

# Apply interpolation only for NA values at Depth_meter = 0.15
EERES_T_long <- EERES_T_long %>%
  group_by(datetime) %>%
  mutate(Water_Temperature_celsius = ifelse(Depth_meter == 0.15 & is.na(Water_Temperature_celsius),
                                             interpolate_na_temperature(cur_data()),
                                             Water_Temperature_celsius)) %>%
  ungroup()

# write Water temperature to csv
write.table(EERES_T_long, file = "LakeEnsemblR_wtemp_profile_EERES.csv", sep = ",", dec = ".", row.names = FALSE, col.names = TRUE, quote = FALSE)
# write.table(EERES_T_dropNAcol, file = "LakeEnsemblR_wtemp_profile_EERES.csv", sep = ",", dec = ".", row.names = FALSE, col.names = TRUE, quote = FALSE)
```

## Meteo Data Preparation

Following code presents the preparation of the meteorological data. This is done by aggregation of the meteorological data to the temporal scale of the observed data (1min to 10min). This is done via averages, and sum of rainfall in mm per hour respectively. LakeEnsemblR requires an input of surface level barometric pressure in Pa, therefore the transformation from hPa was undertaken. The wind direction and speed at 10m above surface should be delivered. The exact location of the anemometer was not known, hence it was assumed that the given data reflects this requirement.

```{r meteoprep}
# Meteorological data
meteo_raw = read_xlsx("Data/20200101-20211231_Wetterstationsdaten_Andreas Lorke.xlsx", 
    skip = 1, na=c("-99999.9","-9999")) 
meteo = meteo_raw %>%
  as.data.frame() %>%
  mutate(across(where(is.numeric), ~ na_if(., -9999))) %>%      
  mutate(across(where(is.numeric), ~ na_if(., -99999.9))) %>%   # refactor NAs
  mutate(time=format(as_datetime(`~...1`),"%H:%M:%S"),
         date=format(as_datetime(`~...2`),"%Y-%m-%d")) %>%                      
  mutate(datetime=ymd(date) + hms(time))                        # datetime from date and time


# check for na values in all columns but the wind direction column WV
na_df <- meteo %>%
  select(SO,TE,FE,TP,WD,WR,RD,DR,RE) %>%
  filter(if_any(everything(), is.na))

# 10min meteo data

meteo_10m_t = meteo %>%
  # Melt the data to long format
  pivot_longer(cols = -c(datetime,`~...1`,`~...2`,time,date), names_to = "variable", values_to = "value") %>%
  # Create time groups
  mutate(time_group = floor_date(datetime, unit = "10 minutes")) %>%
  # Group by time_group and variable
  group_by(time_group, variable) %>%
  # Summarise the data to ensure a single value per group
  summarise(value = ifelse(variable == "RD", sum(value, na.rm = TRUE), mean(value, na.rm = TRUE)), 
            .groups = "drop") %>%
  # Gather back to wide format, ensuring unique values
  pivot_wider(names_from = variable, values_from = value, values_fn = list(value = mean))

# fill missing steps in time series meteo data
complete_time_series <- seq(from = min(meteo_10m_t$time_group, na.rm = TRUE), 
                  to = max(meteo_10m_t$time_group, na.rm = TRUE), 
                  by = "10 min")

meteo_complete <- data.frame(time_group=complete_time_series) %>%
  full_join(meteo_10m_t, by = "time_group")

meteo_filled <- meteo_complete %>%
#interpolate values
  mutate(across(-c(time_group, WV), ~ zoo::na.approx(.))) %>%
  select(time_group,SO,TE,FE,TP,WD,WR,RD,DR) %>%
#hPa to Pa
  mutate(DR = 100*DR) %>%
#mm/10min to mm/hr
  mutate(RD = 6*RD) %>%
#rename and restructure to LakeEnsemblR Standard
  rename(datetime=time_group,
  Shortwave_Radiation_Downwelling_wattPerMeterSquared=SO,
  Air_Temperature_celsius=TE,
  Relative_Humidity_percent=FE,
  Dewpoint_Temperature_celsius=TP,
  Ten_Meter_Elevation_Wind_Speed_meterPerSecond=WD,
  Ten_Meter_Elevation_Wind_Direction_degree=WR,
  Rainfall_millimeterPerHour=RD,
  Surface_Level_Barometric_Pressure_pascal=DR) %>%
  mutate(datetime = format(datetime, "%Y-%m-%d %H:%M:%S")) %>%
  select(datetime,Ten_Meter_Elevation_Wind_Speed_meterPerSecond,Air_Temperature_celsius,Relative_Humidity_percent,Shortwave_Radiation_Downwelling_wattPerMeterSquared,Surface_Level_Barometric_Pressure_pascal,Rainfall_millimeterPerHour)

# check for remaining NAs
meteo_filled %>%
  filter(if_any(everything(), is.na))

# check for missing time steps
all(meteo_filled$time_group == complete_time_series)

write.table(meteo_filled, file = "LakeEnsemblR_meteo_EERES.csv", sep = ",", dec = ".", row.names = FALSE, col.names = TRUE, quote = FALSE)
```

```{r Nullmodeldata, results=hide, include=FALSE}
generate_random_meteo_data <- function(template, n) {
  sample_data <- data.frame(datetime = template$datetime)  # Keep datetime unchanged
  
  # Generate new data for each numeric column
  for (col in names(template)[-1]) {  # Exclude the datetime column
    if (is.numeric(template[[col]])) {
      # Get min and max values for the column
      min_val <- min(template[[col]], na.rm = TRUE)
      max_val <- max(template[[col]], na.rm = TRUE)
      
      # Calculate mean and standard deviation of the original column
      mean_val <- mean(template[[col]], na.rm = TRUE)
      sd_val <- sd(template[[col]], na.rm = TRUE)
      
      # Generate new values based on normal distribution
      new_values <- rnorm(n, mean = mean_val, sd = sd_val)
      
      # Ensure values are within the specified bounds
      new_values <- pmax(new_values, min_val)  # Set a lower bound
      new_values <- pmin(new_values, max_val)  # Set an upper bound
      
      sample_data[[col]] <- new_values
    }
  }
  
  return(sample_data)
}

# Generate a new sample dataframe based on the template
set.seed(123)  # For reproducibility
n_samples <- nrow(meteo_filled)  # Number of samples to generate
new_random_meteo_df <- generate_random_meteo_data(meteo_filled, n_samples)

# Function to generate new random data for Water_Temperature_celsius
generate_random_EERES_data <- function(template, n) {
  sample_data <- data.frame(datetime = template$datetime, Depth_meter = template$Depth_meter)  # Keep datetime and Depth unchanged
  
  # Get min and max values for Water_Temperature_celsius
  min_temp <- min(template$Water_Temperature_celsius, na.rm = TRUE)
  max_temp <- max(template$Water_Temperature_celsius, na.rm = TRUE)
  
  # Generate new temperature values
  mean_temp <- mean(template$Water_Temperature_celsius, na.rm = TRUE)
  sd_temp <- sd(template$Water_Temperature_celsius, na.rm = TRUE)
  
  # Generate new temperature values and ensure they are within the specified bounds
  new_temp_values <- rnorm(n, mean = mean_temp, sd = sd_temp)
  new_temp_values <- pmax(new_temp_values, min_temp)  # Ensure values are not below min_temp
  new_temp_values <- pmin(new_temp_values, max_temp)  # Ensure values are not above max_temp
  
  sample_data$Water_Temperature_celsius <- new_temp_values
  
  return(sample_data)
}


# Generate a new sample dataframe based on the template
set.seed(123)  # For reproducibility
n_samples <- nrow(EERES_T_long)  # Number of samples to generate
new_random_EERES_df <- generate_random_EERES_data(EERES_T_long, n_samples)

write.table(new_random_meteo_df, file = "LakeEnsemblR_meteo.csv", sep = ",", dec = ".", row.names = FALSE, col.names = TRUE, quote = FALSE)

write.table(new_random_EERES_df, file = "LakeEnsemblR_wtemp_profile.csv", sep = ",", dec = ".", row.names = FALSE, col.names = TRUE, quote = FALSE)
```

```{r Nullmodel, results=hide, include=FALSE}
getwd()
# setwd("../")
# Set config file & models
config_file <- "LakeEnsemblR.yaml"
# model <- c("FLake", "GLM", "GOTM", "Simstrat",  "MyLake")
model <- c("FLake","GOTM","Simstrat")

# Example run
# 1. Export settings - creates directories with all model setups and exports settings from the LER configuration file
export_config(config_file = config_file, model = model)

# 2. Run ensemble lake models
run_ensemble(config_file = config_file, model = model, parallel = T, return_list = T, ncores = 4)

ncdf <- 'output/ensemble_output.nc'
vars <- gotmtools::list_vars(ncdf)
vars # Print variables

p1 <- plot_heatmap(ncdf)
p1 <- p1 +
  theme_classic(base_size = 24) + 
  scale_colour_gradientn(limits = c(0, 21),
                         colours = rev(RColorBrewer::brewer.pal(11, "Spectral")))
p1

```

# Model

## Ensemble model run

Following Code runs the LER model. There are issues using GLM and MyLake in shallow ponds, therefore all following simulations are done using exclusively FLake, GOTM or Simstrat. We found that the model FLake calculates a mean depth of the lake from the volume (calculated from truncated square pyramids between the layers) and the surface area. This however seems to reduce the depth of shallow ponds below the depths of observed layers. Therefore we manually set the depth of the pond for FLake using the parameter `depth_w_lk`.

```{r ensemblemodel}
getwd()
# setwd("../")
# Set config file & models
config_file <- "LakeEnsemblR_EERES.yaml"
# model <- c("FLake", "GLM", "GOTM", "Simstrat",  "MyLake")
model <- c("FLake","GOTM","Simstrat")

# Example run
# 1. Export settings - creates directories with all model setups and exports settings from the LER configuration file
export_config(config_file = config_file, model = model)

# 2. Run ensemble lake models
run_ensemble(config_file = config_file, model = model, parallel = T, return_list = T, ncores = 3)

#Check configs
# check_master_config(config_file= config_file, model=model, exp_cnf = T)
```

## Calibration

The calibration is done using the function cali_ensemble. Due to errors by using this function on all models simultaneously, the results will be generated for each model individually For each model, a calibration using a set of 500 parameter combinations is used for the models Simstrat and GOTM. For FLake only 100 parameter combinations were used due to technical obstacles. A Latin Hypercube Calibration (LHC) method is used in the calibration. With the function `plot_LHC`, the results can be shown graphically.

Depending on some arguments given to cali_ensemble (we guess `parallel`), the output of the function will either be the results or the names of the .csv files which store the parameter sets and results. Due to this inconsistency in the code, we manually select the generated results from the folder "cali".

```{r}
# getting the model fit of the last ensemble run
calc_fit(ncdf = "output/ensemble_output.nc", model = model)

config_file = "LakeEnsemblR_EERES_bestparams.yaml"

# # run the calibration using the LHC method
# cali_res_Simstrat <- cali_ensemble(config_file = config_file, num = 500, cmethod = "LHC",
#                           parallel = T, model = "SimStrat", ncores = 3)
# cali_res_GOTM <- cali_ensemble(config_file = config_file, num = 30, cmethod = "LHC",
#                           parallel = T, model = "GOTM", ncores = 3)
cali_res_FLake <- cali_ensemble(config_file = config_file, num = 30, cmethod = "LHC",
                          parallel = T, model = "FLake", ncores = 3)

# we manually set the file names with the results
res_file_GOTM <- paste0(getwd(),
                            c("/cali/params_GOTM_LHC_n500.csv",
                              "/cali/GOTM_LHC_n500.csv"))
res_file_FLake <- paste0(getwd(),
                            c("/cali/params_FLake_LHC_n100.csv",
                              "/cali/FLake_LHC_n100.csv"))
res_file_Simstrat <- paste0(getwd(),
                            c("/cali/params_Simstrat_LHC_n500.csv",
                              "/cali/Simstrat_LHC_n500.csv"))
# # we can plot the results of the latin hypercube calibrations using plot_LHC
# plot_LHC(config_file = config_file, model = "GOTM", res_files = res_file_GOTM,
#          qual_met = "nse", best = "high")
# # or we can plot it for all models using
# plot_LHC(config_file = config_file, model = "FLake", res_files = res_file_FLake,
#          qual_met = "nse", best = "high")
# plot_LHC(config_file = config_file, model = "SimStrat", res_files = res_file_Simstrat,
#          qual_met = "nse", best="high")
```

<!-- The results of the calibration can be extracted using `load_LHC_results`. Subsequently, the set of best parameters using the goodness of model fit measure Root Mean Squared Error (RMSE), Nash-Sutcliffe Efficiency (NSE) and mean absolute error (MAE) is printed. -->

The simulations using different sets of parameters were subsequently analyzed, aiming to identify parameter values which minimize deviations from observed data. For this purpose a composite error metric combining RMSE, NSE, Bias and **R^2** was calculated. Parameter values which minimized this composite score were chosen as the calibrated value.

```{r parameter calibration}
params_Sim = read.csv("C:/Users/euble/OneDrive/Master/ACP5/ACP5_PondModel_Main/ACP5_PondModel_Main/cali/params_Simstrat_LHC_n500.csv")
em_Sim = read.csv("C:/Users/euble/OneDrive/Master/ACP5/ACP5_PondModel_Main/ACP5_PondModel_Main/cali/Simstrat_LHC_n500.csv")

params_GOTM = read.csv("C:/Users/euble/OneDrive/Master/ACP5/ACP5_PondModel_Main/ACP5_PondModel_Main/cali/params_GOTM_LHC_n500.csv")
em_GOTM = read.csv("C:/Users/euble/OneDrive/Master/ACP5/ACP5_PondModel_Main/ACP5_PondModel_Main/cali/GOTM_LHC_n500.csv")

params_FLake = read.csv("C:/Users/euble/OneDrive/Master/ACP5/ACP5_PondModel_Main/ACP5_PondModel_Main/cali/params_FLake_LHC_n100.csv")
em_FLake = read.csv("C:/Users/euble/OneDrive/Master/ACP5/ACP5_PondModel_Main/ACP5_PondModel_Main/cali/FLake_LHC_n100.csv")

join_Sim = full_join(params_Sim, em_Sim, by = "par_id")
join_GOTM = full_join(params_GOTM, em_GOTM, by = "par_id")
join_FLake = full_join(params_FLake, em_FLake, by = "par_id")


## Merge data sets for calibration of shared parameters
Sim_red = join_Sim %>% select(-a_seiche, -par_id)
GOTM_red = join_GOTM %>% select(-turb_param.k_min, -par_id)
FLake_red = join_FLake %>% select(-c_relax_C, -par_id)

shared = rbind(Sim_red, GOTM_red, FLake_red)



#### Calculate composite error metric score ####

## Weights (scaled with standard deviation to approximate equal influence of individual metrics)
w1 = 1/sd(join_Sim$rmse) # RMSE
w2 = 1/sd((1-join_Sim$nse)) # NSE
w3 = 1/sd(abs(join_Sim$bias)) # Bias
w4 = 1/sd(1- join_Sim$r) # R^2

## Compisite error metric
join_Sim$score = w1*join_Sim$rmse + w2*(1-join_Sim$nse) + w3*abs(join_Sim$bias) + w4*(1- join_Sim$r)


## Weights (scaled with standard deviation to approximate equal influence of individual metrics)
w5 = 1/sd(join_GOTM$rmse) # RMSE
w6 = 1/sd((1-join_GOTM$nse)) # NSE
w7 = 1/sd(abs(join_GOTM$bias)) # Bias
w8 = 1/sd(1- join_GOTM$r) # R^2

## Compisite error metric
join_GOTM$score = w5*join_GOTM$rmse + w6*(1-join_GOTM$nse) + w7*abs(join_GOTM$bias) + w8*(1- join_GOTM$r)


## Weights (scaled with standard deviation to approximate equal influence of individual metrics)
w9 = 1/sd(join_FLake$rmse) # RMSE
w10 = 1/sd((1-join_FLake$nse)) # NSE
w11 = 1/sd(abs(join_FLake$bias)) # Bias
w12 = 1/sd(1- join_FLake$r) # R^2

## Compisite error metric
join_FLake$score = w9*join_FLake$rmse + w10*(1-join_FLake$nse) + w11*abs(join_FLake$bias) + w12*(1- join_FLake$r)


## Weights (scaled with standard deviation to approximate equal influence of individual metrics)
w13 = 1/sd(shared$rmse) # RMSE
w14 = 1/sd((1-shared$nse)) # NSE
w15 = 1/sd(abs(shared$bias)) # Bias
w16 = 1/sd(1- shared$r) # R^2

## Compisite error metric
shared$score = w13*shared$rmse + w14*(1-shared$nse) + w15*abs(shared$bias) + w16*(1- shared$r)

##### Shared model parameters #####

## Wind speed

# Extract the computed density data from ggplot
dens_data <- ggplot_build(
  ggplot(shared, aes(x = wind_speed)) +
    geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4)
)$data[[1]]

# Find wind speed where density is maximized
optimal_wind_speed <- dens_data$x[which.max(dens_data$y)]




a_ws = ggplot(shared, aes(x = wind_speed, y = score)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "loess", color = "blue") +
  geom_vline(xintercept = optimal_wind_speed, color = "red", linetype = "dashed") +
  labs(title = "Composite Score vs Wind Speed",
       x = "Wind Speed",
       y = "Composite Score")

b_ws = ggplot(shared, aes(x = wind_speed)) +
  geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4) +
  geom_vline(xintercept = optimal_wind_speed, color = "red", linetype = "dashed") +
  labs(title = "Density of Wind Speed for Low Composite Score",
       x = "Wind Speed",
       y = "Weighted Density")

a_ws+b_ws

## Optimal wind speed
print(optimal_wind_speed)


## SWR
# Extract the computed density data from ggplot
dens_data <- ggplot_build(
  ggplot(shared, aes(x = swr)) +
    geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4)
)$data[[1]]

# Find wind speed where density is maximized
optimal_swr <- dens_data$x[which.max(dens_data$y)]


a_swr = ggplot(shared, aes(x = swr, y = score)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "loess", color = "blue") +
  geom_vline(xintercept = optimal_swr, color = "red", linetype = "dashed") +
  labs(title = "Composite Score vs swr",
       x = "swr",
       y = "Composite Score")

b_swr = ggplot(shared, aes(x = swr)) +
  geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4) +
  geom_vline(xintercept = optimal_swr, color = "red", linetype = "dashed") +
  labs(title = "Density of swr for Low Composite Score",
       x = "swr",
       y = "Weighted Density")

a_swr+b_swr

print(optimal_swr)

## Kw
# Extract the computed density data from ggplot
dens_data <- ggplot_build(
  ggplot(shared, aes(x = Kw)) +
    geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4)
)$data[[1]]

# Find wind speed where density is maximized
optimal_Kw <- dens_data$x[which.max(dens_data$y)]


a_Kw = ggplot(shared, aes(x = Kw, y = score)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "loess", color = "blue") +
  geom_vline(xintercept = optimal_Kw, color = "red", linetype = "dashed") +
  labs(title = "Composite Score vs Kw",
       x = "Kw",
       y = "Composite Score")

b_Kw = ggplot(shared, aes(x = Kw)) +
  geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4) +
  geom_vline(xintercept = optimal_Kw, color = "red", linetype = "dashed") +
  labs(title = "Density of Kw for Low Composite Score",
       x = "Kw",
       y = "Weighted Density")

a_Kw+b_Kw

print(optimal_Kw)


#### Model specific parameters ####

  ##### Simstrat #####
    
    ## a_seiche
    # Extract the computed density data from ggplot
    dens_data <- ggplot_build(
      ggplot(join_Sim, aes(x = a_seiche)) +
        geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4)
    )$data[[1]]
    
    # Find wind speed where density is maximized
    optimal_a_seiche <- dens_data$x[which.max(dens_data$y)]
    
    
    a_a_seiche = ggplot(join_Sim, aes(x = a_seiche, y = score)) +
      geom_point(alpha = 0.5) +
      geom_smooth(method = "loess", color = "blue") +
      geom_vline(xintercept = optimal_a_seiche, color = "red", linetype = "dashed") +
      labs(title = "Composite Score vs a_seiche",
           x = "a_seiche",
           y = "Composite Score")
    
    b_a_seiche = ggplot(join_Sim, aes(x = a_seiche)) +
      geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4) +
      geom_vline(xintercept = optimal_a_seiche, color = "red", linetype = "dashed") +
      labs(title = "Density of a_seiche for Low Composite Score",
           x = "a_seiche",
           y = "Weighted Density")
    
    a_a_seiche+b_a_seiche
    
    print(optimal_a_seiche)
    
  ##### GOTM #####
    
    ## turb_param.k_min
    # Extract the computed density data from ggplot
    dens_data <- ggplot_build(
      ggplot(join_GOTM, aes(x = turb_param.k_min)) +
        geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4)
    )$data[[1]]
    
    # Find wind speed where density is maximized
    optimal_turb_param.k_min <- dens_data$x[which.max(dens_data$y)]
    
    
    a_turb_param.k_min = ggplot(join_GOTM, aes(x = turb_param.k_min, y = score)) +
      geom_point(alpha = 0.5) +
      geom_smooth(method = "loess", color = "blue") +
      geom_vline(xintercept = optimal_turb_param.k_min, color = "red", linetype = "dashed") +
      labs(title = "Composite Score vs turb_param.k_min",
           x = "turb_param.k_min",
           y = "Composite Score")
    
    b_turb_param.k_min = ggplot(join_GOTM, aes(x = turb_param.k_min)) +
      geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4) +
      geom_vline(xintercept = optimal_turb_param.k_min, color = "red", linetype = "dashed") +
      labs(title = "Density of turb_param.k_min for Low Composite Score",
           x = "turb_param.k_min",
           y = "Weighted Density")
    
    a_turb_param.k_min+b_turb_param.k_min
    
    print(optimal_turb_param.k_min)
    
  ##### FLake #####
    
    ## c_relax_C
    # Extract the computed density data from ggplot
    dens_data <- ggplot_build(
      ggplot(join_FLake, aes(x = c_relax_C)) +
        geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4)
    )$data[[1]]
    
    # Find wind speed where density is maximized
    optimal_c_relax_C <- dens_data$x[which.max(dens_data$y)]
    
    
    a_c_relax_C = ggplot(join_FLake, aes(x = c_relax_C, y = score)) +
      geom_point(alpha = 0.5) +
      geom_smooth(method = "loess", color = "blue") +
      geom_vline(xintercept = optimal_c_relax_C, color = "red", linetype = "dashed") +
      labs(title = "Composite Score vs c_relax_C",
           x = "c_relax_C",
           y = "Composite Score")
    
    b_c_relax_C = ggplot(join_FLake, aes(x = c_relax_C)) +
      geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4) +
      geom_vline(xintercept = optimal_c_relax_C, color = "red", linetype = "dashed") +
      labs(title = "Density of c_relax_C for Low Composite Score",
           x = "c_relax_C",
           y = "Weighted Density")
    
    a_c_relax_C+b_c_relax_C
    
    print(optimal_c_relax_C)

```

These optimized parameters were implemented as initial parameters in a new configuration file (LakeEnsemblR_EERES_calibrated.yaml). Note that calibration can be more extensive using different calibration methods and reiterating the simulation and evaluation process. However this would exceed the scope of the project and may lead to overfitting. Nevertheless calibration can be complex problem since different equally valid methods may lead to different calibrated values (See Appendix). The rather simple calibration method shall suffice for the purpose of applying the LakeEnsembleR package in practise.




## Model Run with best parameters

Followingly, the ensemble is run with the calibrated parameters. TODO: all.

```{r bestparamrun}
config_file = "LakeEnsemblR_EERES_calibrated.yaml"
model = c("FLake","GOTM","Simstrat")
export_config(config_file = config_file, model = model)
run_ensemble(config_file = config_file, model = model, parallel=T)
```

# Analysis

Now we plot the heatmap of the ensemble run.

```{r ensembleheatmap}
## Plot model output using gotmtools/ggplot2
# Extract names of all the variables in netCDF
ncdf <- 'output/ensemble_output.nc'
vars <- gotmtools::list_vars(ncdf)
vars # Print variables

p1 <- plot_heatmap(ncdf)
# p1
# Change the theme and increase text size for saving
p1 <- p1 +
  
  theme_classic(base_size = 24) + 
  scale_colour_gradientn(limits = c(0, 21),
                         colours = rev(RColorBrewer::brewer.pal(11, "Spectral")))
# Save as a png file
p1
# ggsave('output/ensemble_heatmap.png', p1,  dpi = 300,width = 384,height = 280, units = 'mm')

```

<<<<<<< HEAD
TODO: not looked into this yet

```{r PlotEnsembleOUptput}
# Plot ensemble mean at 0.9m
model = c("FLake", "GOTM", "Simstrat")
plot_ensemble(ncdf = ncdf, model = model, var = "temp", depth = 0.25)

# Load watertemp from netCDF file as a list
wtemp <- load_var(ncdf = ncdf, var = "temp", return = "list")
names(wtemp)

# Plot residual diagnostic plots
plots <- plot_resid(var_list = wtemp) # Plot residuals - returns a list of plots
names(plots) #

# Plot residuals vs. depth ("res_depth")
plots[[2]]

# Analyse the netCDF output
out <- analyse_ncdf(ncdf, model = model, spin_up = 0)
names(out)
str <- out[["strat"]]
str[str$year == 2010, ]
stats <- out[["stats"]]
stats
```

# Appendix
=======
```{r model results}
## Set up
ncdf <- "output_FLake_GOTM_Simstrat/ensemble_output.nc"
model = c("FLake", "GOTM", "Simstrat")

#### Results ####
# Take a look at the model fits to the observed data
error_metrics = calc_fit(ncdf = ncdf,
         model = model,
         var = "temp")



##### Model fits #####

##### Heatmap #####
p1 <- plot_heatmap(ncdf)

# Change the theme and increase text size for saving
p1 <- p1 +
  theme_classic(base_size = 24) + 
  scale_colour_gradientn(limits = c(0, 21),
                         colours = rev(RColorBrewer::brewer.pal(11, "Spectral")))

p1

##### Built-in: Models vs. Oberservations, Residuals, Temperature range boxplots (per depth) #####

# Plot ensemble mean at 0m
plot_ensemble(ncdf = ncdf, model = model, var = "temp", depth = 0, residuals = TRUE, boxwhisker = TRUE)

# Plot ensemble mean at 0.1m
plot_ensemble(ncdf = ncdf, model = model, var = "temp", depth = 0.1, residuals = TRUE, boxwhisker = TRUE)

# Plot ensemble mean at 0.15m
plot_ensemble(ncdf = ncdf, model = model, var = "temp", depth = 0.15, residuals = TRUE, boxwhisker = TRUE)

# Plot ensemble mean at 0.2m
plot_ensemble(ncdf = ncdf, model = model, var = "temp", depth = 0.2, residuals = TRUE, boxwhisker = TRUE)

# Plot ensemble mean at 0.25m
plot_ensemble(ncdf = ncdf, model = model, var = "temp", depth = 0.25, residuals = TRUE, boxwhisker = TRUE)

# Plot ensemble mean at 0.3m
# plot_ensemble(ncdf = ncdf, model = model, var = "temp", depth = 0.3, residuals = TRUE, boxwhisker = TRUE)



##### ggplot2: Water temperature predictions of models vs. observed temperatures (at depths in m: 0, 0.05, 0.1, 0.15, 0.2, 0.25) #####

# Load post-processed output data into your workspace
analyse_df <- analyse_ncdf(ncdf = ncdf, spin_up = NULL, drho = 0.1, model = model)


# Example plot the summer stratification period
strat_df <- analyse_df$strat
out_df <- analyse_df$out_df
obs <- analyse_df$obs_temp

obs$model = "Observed"
#obs$depths = obs$depths - 0.05
out_df = rbind(out_df, obs)


# Convert date column to POSIXct for proper time handling
out_df <- out_df %>%
  mutate(date = as.POSIXct(date, format = "%Y-%m-%d %H:%M:%S"),
         depth_label = paste0(depths, " m"))  # Create depth label

# Create the plot with labeled facets
ggplot(out_df, aes(x = date, y = temp, color = model)) +
  geom_line() +
  facet_wrap(~depth_label, scales = "free_y") +  # Label depth with "m"
  labs(title = "Water Temperature Predictions by Model",
       x = "Date", y = "Temperature (°C)", color = "Model") +
  theme_minimal() +
  theme(legend.position = "bottom",
        strip.text = element_text(size = 12))  # Make facet labels clearer




```
>>>>>>> 57b123ab197072dcc6eb57472d486dd1525279fa


## Calibration different methods
```{r}

params_Sim = read.csv("C:/Users/euble/OneDrive/Master/ACP5/ACP5_PondModel_Main/ACP5_PondModel_Main/cali/params_Simstrat_LHC_n500.csv")
em_Sim = read.csv("C:/Users/euble/OneDrive/Master/ACP5/ACP5_PondModel_Main/ACP5_PondModel_Main/cali/Simstrat_LHC_n500.csv")

params_GOTM = read.csv("C:/Users/euble/OneDrive/Master/ACP5/ACP5_PondModel_Main/ACP5_PondModel_Main/cali/params_GOTM_LHC_n500.csv")
em_GOTM = read.csv("C:/Users/euble/OneDrive/Master/ACP5/ACP5_PondModel_Main/ACP5_PondModel_Main/cali/GOTM_LHC_n500.csv")

params_FLake = read.csv("C:/Users/euble/OneDrive/Master/ACP5/ACP5_PondModel_Main/ACP5_PondModel_Main/cali/params_FLake_LHC_n100.csv")
em_FLake = read.csv("C:/Users/euble/OneDrive/Master/ACP5/ACP5_PondModel_Main/ACP5_PondModel_Main/cali/FLake_LHC_n100.csv")

join_Sim = full_join(params_Sim, em_Sim, by = "par_id")
join_GOTM = full_join(params_GOTM, em_GOTM, by = "par_id")
join_FLake = full_join(params_FLake, em_FLake, by = "par_id")


## Merge data sets for calibration of shared parameters
Sim_red = join_Sim %>% select(-a_seiche, -par_id)
GOTM_red = join_GOTM %>% select(-turb_param.k_min, -par_id)
FLake_red = join_FLake %>% select(-c_relax_C, -par_id)

shared = rbind(Sim_red, GOTM_red, FLake_red)



#### Calculate composite error metric score ####

  ## Weights (scaled with standard deviation to approximate equal influence of individual metrics)
  w1 = 1/sd(join_Sim$rmse) # RMSE
  w2 = 1/sd((1-join_Sim$nse)) # NSE
  w3 = 1/sd(abs(join_Sim$bias)) # Bias
  w4 = 1/sd(1- join_Sim$r) # R^2
  
  ## Compisite error metric
  join_Sim$score = w1*join_Sim$rmse + w2*(1-join_Sim$nse) + w3*abs(join_Sim$bias) + w4*(1- join_Sim$r)

  
  ## Weights (scaled with standard deviation to approximate equal influence of individual metrics)
  w5 = 1/sd(join_GOTM$rmse) # RMSE
  w6 = 1/sd((1-join_GOTM$nse)) # NSE
  w7 = 1/sd(abs(join_GOTM$bias)) # Bias
  w8 = 1/sd(1- join_GOTM$r) # R^2
  
  ## Compisite error metric
  join_GOTM$score = w5*join_GOTM$rmse + w6*(1-join_GOTM$nse) + w7*abs(join_GOTM$bias) + w8*(1- join_GOTM$r)
  

  ## Weights (scaled with standard deviation to approximate equal influence of individual metrics)
  w9 = 1/sd(join_FLake$rmse) # RMSE
  w10 = 1/sd((1-join_FLake$nse)) # NSE
  w11 = 1/sd(abs(join_FLake$bias)) # Bias
  w12 = 1/sd(1- join_FLake$r) # R^2
  
  ## Compisite error metric
  join_FLake$score = w9*join_FLake$rmse + w10*(1-join_FLake$nse) + w11*abs(join_FLake$bias) + w12*(1- join_FLake$r)
  
  
  ## Weights (scaled with standard deviation to approximate equal influence of individual metrics)
  w13 = 1/sd(shared$rmse) # RMSE
  w14 = 1/sd((1-shared$nse)) # NSE
  w15 = 1/sd(abs(shared$bias)) # Bias
  w16 = 1/sd(1- shared$r) # R^2
  
  ## Compisite error metric
  shared$score = w13*shared$rmse + w14*(1-shared$nse) + w15*abs(shared$bias) + w16*(1- shared$r)
  
  
  
 
#### Simstrat ####
##### Visual assessment #####

  ## Wind speed

  # Extract the computed density data from ggplot
  dens_data <- ggplot_build(
    ggplot(join_Sim, aes(x = wind_speed)) +
      geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4)
  )$data[[1]]
  
  # Find wind speed where density is maximized
  optimal_wind_speed <- dens_data$x[which.max(dens_data$y)]
  
  
  
  
  a_ws = ggplot(join_Sim, aes(x = wind_speed, y = score)) +
    geom_point(alpha = 0.5) +
    geom_smooth(method = "loess", color = "blue") +
    geom_vline(xintercept = optimal_wind_speed, color = "red", linetype = "dashed") +
    labs(title = "Composite Score vs Wind Speed",
         x = "Wind Speed",
         y = "Composite Score")
  
  b_ws = ggplot(join_Sim, aes(x = wind_speed)) +
    geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4) +
    geom_vline(xintercept = optimal_wind_speed, color = "red", linetype = "dashed") +
    labs(title = "Density of Wind Speed for Low Composite Score",
         x = "Wind Speed",
         y = "Weighted Density")
  
  a_ws+b_ws
  
  ## Optimal wind speed
  print(optimal_wind_speed)
  
  
  ## SWR
  # Extract the computed density data from ggplot
  dens_data <- ggplot_build(
    ggplot(join_Sim, aes(x = swr)) +
      geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4)
  )$data[[1]]
  
  # Find wind speed where density is maximized
  optimal_swr <- dens_data$x[which.max(dens_data$y)]
  
  
  a_swr = ggplot(join_Sim, aes(x = swr, y = score)) +
    geom_point(alpha = 0.5) +
    geom_smooth(method = "loess", color = "blue") +
    geom_vline(xintercept = optimal_swr, color = "red", linetype = "dashed") +
    labs(title = "Composite Score vs swr",
         x = "swr",
         y = "Composite Score")
  
  b_swr = ggplot(join_Sim, aes(x = swr)) +
    geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4) +
    geom_vline(xintercept = optimal_swr, color = "red", linetype = "dashed") +
    labs(title = "Density of swr for Low Composite Score",
         x = "swr",
         y = "Weighted Density")
  
  a_swr+b_swr
  
  print(optimal_swr)
  
  ## Kw
  # Extract the computed density data from ggplot
  dens_data <- ggplot_build(
    ggplot(join_Sim, aes(x = Kw)) +
      geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4)
  )$data[[1]]
  
  # Find wind speed where density is maximized
  optimal_Kw <- dens_data$x[which.max(dens_data$y)]
  
  
  a_Kw = ggplot(join_Sim, aes(x = Kw, y = score)) +
    geom_point(alpha = 0.5) +
    geom_smooth(method = "loess", color = "blue") +
    geom_vline(xintercept = optimal_Kw, color = "red", linetype = "dashed") +
    labs(title = "Composite Score vs Kw",
         x = "Kw",
         y = "Composite Score")
  
  b_Kw = ggplot(join_Sim, aes(x = Kw)) +
    geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4) +
    geom_vline(xintercept = optimal_Kw, color = "red", linetype = "dashed") +
    labs(title = "Density of Kw for Low Composite Score",
         x = "Kw",
         y = "Weighted Density")
  
  a_Kw+b_Kw
  
  print(optimal_Kw)
  
  ## a_seiche
  # Extract the computed density data from ggplot
  dens_data <- ggplot_build(
    ggplot(join_Sim, aes(x = a_seiche)) +
      geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4)
  )$data[[1]]
  
  # Find wind speed where density is maximized
  optimal_a_seiche <- dens_data$x[which.max(dens_data$y)]
  
  
  a_a_seiche = ggplot(join_Sim, aes(x = a_seiche, y = score)) +
    geom_point(alpha = 0.5) +
    geom_smooth(method = "loess", color = "blue") +
    geom_vline(xintercept = optimal_a_seiche, color = "red", linetype = "dashed") +
    labs(title = "Composite Score vs a_seiche",
         x = "a_seiche",
         y = "Composite Score")
  
  b_a_seiche = ggplot(join_Sim, aes(x = a_seiche)) +
    geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4) +
    geom_vline(xintercept = optimal_a_seiche, color = "red", linetype = "dashed") +
    labs(title = "Density of a_seiche for Low Composite Score",
         x = "a_seiche",
         y = "Weighted Density")
  
  a_a_seiche+b_a_seiche
  
  print(optimal_a_seiche)
  
  
  ##### Random Forest predictions #####
  ## Use a random forest model to predict the best set of parameters (using brute force grid search) 
  rf_model <- randomForest(score ~ wind_speed + swr + Kw + a_seiche, 
                           data = join_Sim, 
                           ntree = 500, importance = TRUE)
  
  print(rf_model) # ~82% of variance explained
  
  
  importance(rf_model)  
  varImpPlot(rf_model)  # Plot variable importance
  
  
  # Define ranges of parameters
  wind_range <- seq(min(join_Sim$wind_speed), max(join_Sim$wind_speed), length.out = 50)
  swr_range <- seq(min(join_Sim$swr), max(join_Sim$swr), length.out = 50)
  Kw_range <- seq(min(join_Sim$Kw), max(join_Sim$Kw), length.out = 50)
  a_seiche_range <- seq(min(join_Sim$a_seiche), max(join_Sim$a_seiche), length.out = 50)
  
  # Create all combinations
  param_grid <- expand.grid(wind_speed = wind_range, swr = swr_range, Kw = Kw_range, a_seiche = a_seiche_range)
  
  # Predict scores
  param_grid$predicted_score <- predict(rf_model, newdata = param_grid)
  
  # Determine the 10% quantile threshold
  quantile_10 <- quantile(param_grid$predicted_score, probs = 0.0001)
  
  # Filter the dataset for the lowest 10% of predicted_score values
  filtered_data <- param_grid[param_grid$predicted_score <= quantile_10, ]
  
  # Calculate the mean for each parameter in the filtered data
  mean_values <- colMeans(filtered_data[, c("wind_speed", "swr", "Kw", "a_seiche")])
  
  
  # Print the results
  print(mean_values)
  
  

  

  ##### Rank #####
  ## Ranking parameters sets on composite error score and describing parameter ranges of best sets
  
  ## Finding optimal parameters by ranking parameter sets by composite error score
  join_Sim$rank <- rank(join_Sim$score)
  
  # Select the best 20% of models
  best_params_rank <- join_Sim[join_Sim$rank <= quantile(join_Sim$rank, 0.1), ]
  
  # Summarize the best parameter ranges
  #summary(best_params_rank[, c("wind_speed", "swr", "Kw", "a_seiche")])
  rank = sapply((best_params_rank[, c("wind_speed", "swr", "Kw", "a_seiche")]), mean)
  
  
  #### Comparing results
  density_Simstrat = c(optimal_wind_speed, optimal_swr, optimal_Kw, optimal_a_seiche, "Optimal density")
  rf_Simstrat = c(mean_values, "Random Forest")
  rank_Simstrat = c(rank, "Rank")  
  
  comparison_Simstrat = rbind(density_Simstrat, rf_Simstrat, rank_Simstrat)
  comparison_Simstrat = as.data.frame(comparison_Simstrat)
  names(comparison_Simstrat) = c("wind_speed", "swr", "Kw", "a_seiche", "Calibration method")
  
  
  
  
  
#### GOTM ####
  ##### Visual assessment #####
  
  ## Wind speed
  
  # Extract the computed density data from ggplot
  dens_data <- ggplot_build(
    ggplot(join_GOTM, aes(x = wind_speed)) +
      geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4)
  )$data[[1]]
  
  # Find wind speed where density is maximized
  optimal_wind_speed <- dens_data$x[which.max(dens_data$y)]
  
  
  
  
  a_ws = ggplot(join_GOTM, aes(x = wind_speed, y = score)) +
    geom_point(alpha = 0.5) +
    geom_smooth(method = "loess", color = "blue") +
    geom_vline(xintercept = optimal_wind_speed, color = "red", linetype = "dashed") +
    labs(title = "Composite Score vs Wind Speed",
         x = "Wind Speed",
         y = "Composite Score")
  
  b_ws = ggplot(join_GOTM, aes(x = wind_speed)) +
    geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4) +
    geom_vline(xintercept = optimal_wind_speed, color = "red", linetype = "dashed") +
    labs(title = "Density of Wind Speed for Low Composite Score",
         x = "Wind Speed",
         y = "Weighted Density")
  
  a_ws+b_ws
  
  ## Optimal wind speed
  print(optimal_wind_speed)
  
  
  ## SWR
  # Extract the computed density data from ggplot
  dens_data <- ggplot_build(
    ggplot(join_GOTM, aes(x = swr)) +
      geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4)
  )$data[[1]]
  
  # Find wind speed where density is maximized
  optimal_swr <- dens_data$x[which.max(dens_data$y)]
  
  
  a_swr = ggplot(join_GOTM, aes(x = swr, y = score)) +
    geom_point(alpha = 0.5) +
    geom_smooth(method = "loess", color = "blue") +
    geom_vline(xintercept = optimal_swr, color = "red", linetype = "dashed") +
    labs(title = "Composite Score vs swr",
         x = "swr",
         y = "Composite Score")
  
  b_swr = ggplot(join_GOTM, aes(x = swr)) +
    geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4) +
    geom_vline(xintercept = optimal_swr, color = "red", linetype = "dashed") +
    labs(title = "Density of swr for Low Composite Score",
         x = "swr",
         y = "Weighted Density")
  
  a_swr+b_swr
  
  print(optimal_swr)
  
  ## Kw
  # Extract the computed density data from ggplot
  dens_data <- ggplot_build(
    ggplot(join_GOTM, aes(x = Kw)) +
      geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4)
  )$data[[1]]
  
  # Find wind speed where density is maximized
  optimal_Kw <- dens_data$x[which.max(dens_data$y)]
  
  
  a_Kw = ggplot(join_GOTM, aes(x = Kw, y = score)) +
    geom_point(alpha = 0.5) +
    geom_smooth(method = "loess", color = "blue") +
    geom_vline(xintercept = optimal_Kw, color = "red", linetype = "dashed") +
    labs(title = "Composite Score vs Kw",
         x = "Kw",
         y = "Composite Score")
  
  b_Kw = ggplot(join_GOTM, aes(x = Kw)) +
    geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4) +
    geom_vline(xintercept = optimal_Kw, color = "red", linetype = "dashed") +
    labs(title = "Density of Kw for Low Composite Score",
         x = "Kw",
         y = "Weighted Density")
  
  a_Kw+b_Kw
  
  print(optimal_Kw)
  
  ## turb_param.k_min
  # Extract the computed density data from ggplot
  dens_data <- ggplot_build(
    ggplot(join_GOTM, aes(x = turb_param.k_min)) +
      geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4)
  )$data[[1]]
  
  # Find wind speed where density is maximized
  optimal_turb_param.k_min <- dens_data$x[which.max(dens_data$y)]
  
  
  a_turb_param.k_min = ggplot(join_GOTM, aes(x = turb_param.k_min, y = score)) +
    geom_point(alpha = 0.5) +
    geom_smooth(method = "loess", color = "blue") +
    geom_vline(xintercept = optimal_turb_param.k_min, color = "red", linetype = "dashed") +
    labs(title = "Composite Score vs turb_param.k_min",
         x = "turb_param.k_min",
         y = "Composite Score")
  
  b_turb_param.k_min = ggplot(join_GOTM, aes(x = turb_param.k_min)) +
    geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4) +
    geom_vline(xintercept = optimal_turb_param.k_min, color = "red", linetype = "dashed") +
    labs(title = "Density of turb_param.k_min for Low Composite Score",
         x = "turb_param.k_min",
         y = "Weighted Density")
  
  a_turb_param.k_min+b_turb_param.k_min
  
  print(optimal_turb_param.k_min)
  
  
  ##### Random Forest predictions #####
  ## Use a random forest model to predict the best set of parameters (using brute force grid search) 
  rf_model <- randomForest(score ~ wind_speed + swr + Kw + turb_param.k_min, 
                           data = join_GOTM, 
                           ntree = 500, importance = TRUE)
  
  print(rf_model) # ~48% of variance explained

  importance(rf_model)  
  varImpPlot(rf_model)  # Plot variable importance
  
  
  # Define ranges of parameters
  wind_range <- seq(min(join_GOTM$wind_speed), max(join_GOTM$wind_speed), length.out = 50)
  swr_range <- seq(min(join_GOTM$swr), max(join_GOTM$swr), length.out = 50)
  Kw_range <- seq(min(join_GOTM$Kw), max(join_GOTM$Kw), length.out = 50)
  turb_param.k_min_range <- seq(min(join_GOTM$turb_param.k_min), max(join_GOTM$turb_param.k_min), length.out = 50)
  
  # Create all combinations
  param_grid <- expand.grid(wind_speed = wind_range, swr = swr_range, Kw = Kw_range, turb_param.k_min = turb_param.k_min_range)
  
  # Predict scores
  param_grid$predicted_score <- predict(rf_model, newdata = param_grid)
  
  # Determine the 0.01% quantile threshold
  quantile_10 <- quantile(param_grid$predicted_score, probs = 0.0001)
  
  # Filter the dataset for the lowest 10% of predicted_score values
  filtered_data <- param_grid[param_grid$predicted_score <= quantile_10, ]
  
  # Calculate the mean for each parameter in the filtered data
  mean_values <- colMeans(filtered_data[, c("wind_speed", "swr", "Kw", "turb_param.k_min")])
  
  # Print the results
  print(mean_values)
  
  
  
  
  
  ##### Rank #####
  ## Ranking parameters sets on composite error score and describing parameter ranges of best sets
  
  ## Finding optimal parameters by ranking parameter sets by composite error score
  join_GOTM$rank <- rank(join_GOTM$score)
  
  # Select the best 20% of models
  best_params_rank <- join_GOTM[join_GOTM$rank <= quantile(join_GOTM$rank, 0.1), ]
  
  # Summarize the best parameter ranges
  #summary(best_params_rank[, c("wind_speed", "swr", "Kw", "turb_param.k_min")])
  rank = sapply((best_params_rank[, c("wind_speed", "swr", "Kw", "turb_param.k_min")]), mean)
  
  
  #### Comparing results
  density_GOTM = c(optimal_wind_speed, optimal_swr, optimal_Kw, optimal_turb_param.k_min, "Optimal density")
  rf_GOTM = c(mean_values, "Random Forest")
  rank_GOTM = c(rank, "Rank")  
  
  comparison_GOTM = rbind(density_GOTM, rf_GOTM, rank_GOTM)
  comparison_GOTM = as.data.frame(comparison_GOTM)
  names(comparison_GOTM) = c("wind_speed", "swr", "Kw", "turb_param.k_min", "Calibration method")
  
  
  
  
  
  
  
  
#### FLake ####
  ##### Visual assessment #####
  
  ## Wind speed
  
  # Extract the computed density data from ggplot
  dens_data <- ggplot_build(
    ggplot(join_FLake, aes(x = wind_speed)) +
      geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4)
  )$data[[1]]
  
  # Find wind speed where density is maximized
  optimal_wind_speed <- dens_data$x[which.max(dens_data$y)]
  
  
  
  
  a_ws = ggplot(join_FLake, aes(x = wind_speed, y = score)) +
    geom_point(alpha = 0.5) +
    geom_smooth(method = "loess", color = "blue") +
    geom_vline(xintercept = optimal_wind_speed, color = "red", linetype = "dashed") +
    labs(title = "Composite Score vs Wind Speed",
         x = "Wind Speed",
         y = "Composite Score")
  
  b_ws = ggplot(join_FLake, aes(x = wind_speed)) +
    geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4) +
    geom_vline(xintercept = optimal_wind_speed, color = "red", linetype = "dashed") +
    labs(title = "Density of Wind Speed for Low Composite Score",
         x = "Wind Speed",
         y = "Weighted Density")
  
  a_ws+b_ws
  
  ## Optimal wind speed
  print(optimal_wind_speed)
  
  
  ## SWR
  # Extract the computed density data from ggplot
  dens_data <- ggplot_build(
    ggplot(join_FLake, aes(x = swr)) +
      geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4)
  )$data[[1]]
  
  # Find wind speed where density is maximized
  optimal_swr <- dens_data$x[which.max(dens_data$y)]
  
  
  a_swr = ggplot(join_FLake, aes(x = swr, y = score)) +
    geom_point(alpha = 0.5) +
    geom_smooth(method = "loess", color = "blue") +
    geom_vline(xintercept = optimal_swr, color = "red", linetype = "dashed") +
    labs(title = "Composite Score vs swr",
         x = "swr",
         y = "Composite Score")
  
  b_swr = ggplot(join_FLake, aes(x = swr)) +
    geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4) +
    geom_vline(xintercept = optimal_swr, color = "red", linetype = "dashed") +
    labs(title = "Density of swr for Low Composite Score",
         x = "swr",
         y = "Weighted Density")
  
  a_swr+b_swr
  
  print(optimal_swr)
  
  ## Kw
  # Extract the computed density data from ggplot
  dens_data <- ggplot_build(
    ggplot(join_FLake, aes(x = Kw)) +
      geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4)
  )$data[[1]]
  
  # Find wind speed where density is maximized
  optimal_Kw <- dens_data$x[which.max(dens_data$y)]
  
  
  a_Kw = ggplot(join_FLake, aes(x = Kw, y = score)) +
    geom_point(alpha = 0.5) +
    geom_smooth(method = "loess", color = "blue") +
    geom_vline(xintercept = optimal_Kw, color = "red", linetype = "dashed") +
    labs(title = "Composite Score vs Kw",
         x = "Kw",
         y = "Composite Score")
  
  b_Kw = ggplot(join_FLake, aes(x = Kw)) +
    geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4) +
    geom_vline(xintercept = optimal_Kw, color = "red", linetype = "dashed") +
    labs(title = "Density of Kw for Low Composite Score",
         x = "Kw",
         y = "Weighted Density")
  
  a_Kw+b_Kw
  
  print(optimal_Kw)
  
  ## c_relax_C
  # Extract the computed density data from ggplot
  dens_data <- ggplot_build(
    ggplot(join_FLake, aes(x = c_relax_C)) +
      geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4)
  )$data[[1]]
  
  # Find wind speed where density is maximized
  optimal_c_relax_C <- dens_data$x[which.max(dens_data$y)]
  
  
  a_c_relax_C = ggplot(join_FLake, aes(x = c_relax_C, y = score)) +
    geom_point(alpha = 0.5) +
    geom_smooth(method = "loess", color = "blue") +
    geom_vline(xintercept = optimal_c_relax_C, color = "red", linetype = "dashed") +
    labs(title = "Composite Score vs c_relax_C",
         x = "c_relax_C",
         y = "Composite Score")
  
  b_c_relax_C = ggplot(join_FLake, aes(x = c_relax_C)) +
    geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4) +
    geom_vline(xintercept = optimal_c_relax_C, color = "red", linetype = "dashed") +
    labs(title = "Density of c_relax_C for Low Composite Score",
         x = "c_relax_C",
         y = "Weighted Density")
  
  a_c_relax_C+b_c_relax_C
  
  print(optimal_c_relax_C)
  
  
  ##### Random Forest predictions #####
  ## Use a random forest model to predict the best set of parameters (using brute force grid search) 
  rf_model <- randomForest(score ~ wind_speed + swr + Kw + c_relax_C, 
                           data = join_FLake, 
                           ntree = 500, importance = TRUE)
  
  importance(rf_model)  
  varImpPlot(rf_model)  # Plot variable importance
  
  print(rf_model) # ~80% of variance explained
  
  # Define ranges of parameters
  wind_range <- seq(min(join_FLake$wind_speed), max(join_FLake$wind_speed), length.out = 50)
  swr_range <- seq(min(join_FLake$swr), max(join_FLake$swr), length.out = 50)
  Kw_range <- seq(min(join_FLake$Kw), max(join_FLake$Kw), length.out = 50)
  c_relax_C_range <- seq(min(join_FLake$c_relax_C), max(join_FLake$c_relax_C), length.out = 50)
  
  # Create all combinations
  param_grid <- expand.grid(wind_speed = wind_range, swr = swr_range, Kw = Kw_range, c_relax_C = c_relax_C_range)
  
  # Predict scores
  param_grid$predicted_score <- predict(rf_model, newdata = param_grid)
  
  # Determine the 0.01% quantile threshold
  quantile_10 <- quantile(param_grid$predicted_score, probs = 0.0001)
  
  # Filter the dataset for the lowest 10% of predicted_score values
  filtered_data <- param_grid[param_grid$predicted_score <= quantile_10, ]
  
  # Calculate the mean for each parameter in the filtered data
  mean_values <- colMeans(filtered_data[, c("wind_speed", "swr", "Kw", "c_relax_C")])
  
  # Print the results
  print(mean_values)
  
  
  
  
  
  ##### Rank #####
  ## Ranking parameters sets on composite error score and describing parameter ranges of best sets
  
  ## Finding optimal parameters by ranking parameter sets by composite error score
  join_FLake$rank <- rank(join_FLake$score)
  
  # Select the best 20% of models
  best_params_rank <- join_FLake[join_FLake$rank <= quantile(join_FLake$rank, 0.1), ]
  
  # Summarize the best parameter ranges
  #summary(best_params_rank[, c("wind_speed", "swr", "Kw", "c_relax_C")])
  rank = sapply((best_params_rank[, c("wind_speed", "swr", "Kw", "c_relax_C")]), mean)
  
  
  #### Comparing results
  density_FLake = c(optimal_wind_speed, optimal_swr, optimal_Kw, optimal_c_relax_C, "Optimal density")
  rf_FLake = c(mean_values, "Random Forest")
  rank_FLake = c(rank, "Rank")  
  
  comparison_FLake = rbind(density_FLake, rf_FLake, rank_FLake)
  comparison_FLake = as.data.frame(comparison_FLake)
  names(comparison_FLake) = c("wind_speed", "swr", "Kw", "c_relax_C", "Calibration method")
  
  
  
#### Shared parameters ####
  ##### Visual assessment #####
  
  ## Wind speed
  
  # Extract the computed density data from ggplot
  dens_data <- ggplot_build(
    ggplot(shared, aes(x = wind_speed)) +
      geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4)
  )$data[[1]]
  
  # Find wind speed where density is maximized
  optimal_wind_speed <- dens_data$x[which.max(dens_data$y)]
  
  
  
  
  a_ws = ggplot(shared, aes(x = wind_speed, y = score)) +
    geom_point(alpha = 0.5) +
    geom_smooth(method = "loess", color = "blue") +
    geom_vline(xintercept = optimal_wind_speed, color = "red", linetype = "dashed") +
    labs(title = "Composite Score vs Wind Speed",
         x = "Wind Speed",
         y = "Composite Score")
  
  b_ws = ggplot(shared, aes(x = wind_speed)) +
    geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4) +
    geom_vline(xintercept = optimal_wind_speed, color = "red", linetype = "dashed") +
    labs(title = "Density of Wind Speed for Low Composite Score",
         x = "Wind Speed",
         y = "Weighted Density")
  
  a_ws+b_ws
  
  ## Optimal wind speed
  print(optimal_wind_speed)
  
  
  ## SWR
  # Extract the computed density data from ggplot
  dens_data <- ggplot_build(
    ggplot(shared, aes(x = swr)) +
      geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4)
  )$data[[1]]
  
  # Find wind speed where density is maximized
  optimal_swr <- dens_data$x[which.max(dens_data$y)]
  
  
  a_swr = ggplot(shared, aes(x = swr, y = score)) +
    geom_point(alpha = 0.5) +
    geom_smooth(method = "loess", color = "blue") +
    geom_vline(xintercept = optimal_swr, color = "red", linetype = "dashed") +
    labs(title = "Composite Score vs swr",
         x = "swr",
         y = "Composite Score")
  
  b_swr = ggplot(shared, aes(x = swr)) +
    geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4) +
    geom_vline(xintercept = optimal_swr, color = "red", linetype = "dashed") +
    labs(title = "Density of swr for Low Composite Score",
         x = "swr",
         y = "Weighted Density")
  
  a_swr+b_swr
  
  print(optimal_swr)
  
  ## Kw
  # Extract the computed density data from ggplot
  dens_data <- ggplot_build(
    ggplot(shared, aes(x = Kw)) +
      geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4)
  )$data[[1]]
  
  # Find wind speed where density is maximized
  optimal_Kw <- dens_data$x[which.max(dens_data$y)]
  
  
  a_Kw = ggplot(shared, aes(x = Kw, y = score)) +
    geom_point(alpha = 0.5) +
    geom_smooth(method = "loess", color = "blue") +
    geom_vline(xintercept = optimal_Kw, color = "red", linetype = "dashed") +
    labs(title = "Composite Score vs Kw",
         x = "Kw",
         y = "Composite Score")
  
  b_Kw = ggplot(shared, aes(x = Kw)) +
    geom_density(aes(weight = 1 / score), fill = "blue", alpha = 0.4) +
    geom_vline(xintercept = optimal_Kw, color = "red", linetype = "dashed") +
    labs(title = "Density of Kw for Low Composite Score",
         x = "Kw",
         y = "Weighted Density")
  
  a_Kw+b_Kw
  
  print(optimal_Kw)
  
 
  
  
  ##### Random Forest predictions #####
  ## Use a random forest model to predict the best set of parameters (using brute force grid search) 
  rf_model <- randomForest(score ~ wind_speed + swr + Kw, 
                           data = shared, 
                           ntree = 500, importance = TRUE)
  
  importance(rf_model)  
  varImpPlot(rf_model)  # Plot variable importance
  
  print(rf_model) # ~57% of variance explained
  
  # Define ranges of parameters
  wind_range <- seq(min(shared$wind_speed), max(shared$wind_speed), length.out = 50)
  swr_range <- seq(min(shared$swr), max(shared$swr), length.out = 50)
  Kw_range <- seq(min(shared$Kw), max(shared$Kw), length.out = 50)

  # Create all combinations
  param_grid <- expand.grid(wind_speed = wind_range, swr = swr_range, Kw = Kw_range)
  
  # Predict scores
  param_grid$predicted_score <- predict(rf_model, newdata = param_grid)
  
  # Determine the 0.1% quantile threshold
  quantile_10 <- quantile(param_grid$predicted_score, probs = 0.001)
  
  # Filter the dataset for the lowest 10% of predicted_score values
  filtered_data <- param_grid[param_grid$predicted_score <= quantile_10, ]
  
  # Calculate the mean for each parameter in the filtered data
  mean_values <- colMeans(filtered_data[, c("wind_speed", "swr", "Kw")])
  
  # Print the results
  print(mean_values)
  
  
  
  
  
  ##### Rank #####
  ## Ranking parameters sets on composite error score and describing parameter ranges of best sets
  
  ## Finding optimal parameters by ranking parameter sets by composite error score
  shared$rank <- rank(shared$score)
  
  # Select the best 20% of models
  best_params_rank <- shared[shared$rank <= quantile(shared$rank, 0.1), ]
  
  # Summarize the best parameter ranges
  #summary(best_params_rank[, c("wind_speed", "swr", "Kw", "c_relax_C")])
  rank = sapply((best_params_rank[, c("wind_speed", "swr", "Kw")]), mean)
  
  
  #### Comparing results
  density_shared = c(optimal_wind_speed, optimal_swr, optimal_Kw, "Optimal density")
  rf_shared = c(mean_values, "Random Forest")
  rank_shared = c(rank, "Rank")  
  
  comparison_shared = rbind(density_shared, rf_shared, rank_shared)
  comparison_shared = as.data.frame(comparison_shared)
  names(comparison_shared) = c("wind_speed", "swr", "Kw", "Calibration method")
  
  
  #### Visualisation ####
  rf_Simstrat = as.numeric(rf_Simstrat[-c(5)])
  rf_GOTM = as.numeric(rf_GOTM[-c(5)])
  rf_FLake = as.numeric(rf_Simstrat[-c(5)])
  rf_shared = as.numeric(rf_shared[-c(4)])
  
  rank_Simstrat = as.numeric(rank_Simstrat[-c(5)])
  rank_GOTM = as.numeric(rank_GOTM[-c(5)])
  rank_FLake = as.numeric(rank_Simstrat[-c(5)])
  rank_shared = as.numeric(rank_shared[-c(4)])
  
  overview <- data.frame(
    
    value =   as.numeric(c(density_Simstrat[-c(5)], density_GOTM[-c(5)], density_FLake[-c(5)], density_shared[-c(4)],
                rf_Simstrat, rf_GOTM, rf_FLake, rf_shared,
                rank_Simstrat, rank_GOTM, rank_FLake, rank_shared)),
  
    parameter = c("wind_speed", "swr", "Kw", "a_seiche",
                  "wind_speed", "swr", "Kw", "turb_param.k_min",
                  "wind_speed", "swr", "Kw", "c_relax_C",
                  "wind_speed", "swr", "Kw",
                  "wind_speed", "swr", "Kw", "a_seiche",
                  "wind_speed", "swr", "Kw", "turb_param.k_min",
                  "wind_speed", "swr", "Kw", "c_relax_C",
                  "wind_speed", "swr", "Kw",
                  "wind_speed", "swr", "Kw", "a_seiche",
                  "wind_speed", "swr", "Kw", "turb_param.k_min",
                  "wind_speed", "swr", "Kw", "c_relax_C",
                  "wind_speed", "swr", "Kw"),
  
    model = c(rep("Simstrat", 4), rep("GOTM", 4), rep("FLake", 4), rep("Shared", 3),
              rep("Simstrat", 4), rep("GOTM", 4), rep("FLake", 4), rep("Shared", 3),
              rep("Simstrat", 4), rep("GOTM", 4), rep("FLake", 4), rep("Shared", 3)),
    calibration_method = c(rep("Optimal density", 15), rep("Random Forest", 15), rep("Rank", 15)),
    stringsAsFactors = FALSE
  )
  
  
 
  ## parameter values per model and calibration method
  
    ## Shared parameters
    ggplot(overview[overview$parameter %in% c("Kw", "swr", "wind_speed"),], aes(x = model, y = value, fill = calibration_method)) +
      geom_bar(stat = "identity", position = "dodge") +
      facet_wrap(~ parameter, scales = "free_y") +  # One plot per parameter
      labs(title = "Parameter Values by Model and Calibration Method",
           x = "Model",
           y = "Value") +
      theme_minimal()
  
    ## Unique parameters
    ggplot(overview[overview$parameter %in% c("a_seiche", "turb_param.k_min", "c_relax_C"),], aes(x = model, y = value, fill = calibration_method)) +
      geom_bar(stat = "identity", position = "dodge") +
      facet_wrap(~ parameter, scales = "free_y") +  # One plot per parameter
      labs(title = "Parameter Values by Model and Calibration Method",
           x = "Model",
           y = "Value") +
      theme_minimal()
    
    
  ## Distribution of parameter values
    
    ## Shared
    ggplot(overview[overview$parameter %in% c("Kw", "swr", "wind_speed"),], aes(x = parameter, y = value)) +
      geom_boxplot() + geom_jitter() +
      labs(title = "Distribution of Parameter Values Across Calibration Methods",
           x = "Parameter",
           y = "Value") +
      theme_minimal()
  
    ## Unique
    ggplot(overview[overview$parameter %in% c("a_seiche", "turb_param.k_min", "c_relax_C"),], aes(x = parameter, y = value)) +
      geom_boxplot() + geom_jitter() +
    labs(title = "Distribution of Parameter Values Across Calibration Methods",
         x = "Parameter",
         y = "Value") +
      theme_minimal()
    
    
```

## Single model runs

Following Code can be used to run single models and check for errors.

```{r singlemodelruns}
setwd("C:/Users/morit/Documents/ACP5_PondModel")
## Test FLake ----
# 1. Example - creates directories with all model setup
export_config(config_file = config_file, model = "FLake")

setwd("FLake")

FLakeR::run_flake(verbose = TRUE)
out <- read.table("output/output.dat", skip = 1, header = TRUE)
plot(out$time, out$Ts, type = "l")

setwd("../")

## Test GLM ----
# 1. Example - creates directories with all model setup
export_config(config_file = config_file, model = "GLM")

setwd("GLM")

GLM3r::run_glm(verbose = TRUE)
out <- glmtools::get_var("output/output.nc", "temp", reference = "surface")
plot(out$DateTime, out[, 2], type = "l")

setwd("../")

## Test GOTM ----
# 1. Example - creates directories with all model setup
export_config(config_file = config_file, model = "GOTM")

setwd("GOTM")

GOTMr::run_gotm(verbose = TRUE)
out <- gotmtools::get_vari("output/output.nc", "temp")
plot(out$time, out[, 2], type = "l")

setwd("../")

## Test Simstrat ----
# 1. Example - creates directories with all model setup
export_config(config_file = config_file, model = "Simstrat")

setwd("Simstrat")

SimstratR::run_simstrat(par_file = "simstrat.par", verbose = TRUE)
out <- read.table(file.path("output", "T_out.dat"), header = TRUE, 
                  sep = ",", check.names = FALSE)
plot(1:nrow(out), out[, ncol(out)], type = "l")

setwd("../")


## Test MyLake ----
# 1. Example - creates directories with all model setup
export_config(config_file = config_file, model = "MyLake")

# setwd("MyLake")
MyLakeR::run_mylake(sim_folder = ".", config_dat = "mylake.Rdata")
load("/MyLake/output/output.RData")
plot(res$tt, res$Tzt[1, ], type = "l")

setwd("../")

# END
```

## Model runs of enlarged basin

TODO: Think about setting fix_wtlvl to false since this may be important for stratification dynamics. TODO: Set the initial water temperature on a scientifically based information (Use the known temperature and add lower temperatures at lower depths)

```{r scaledmodel}

# 10x depth ----

getwd()

config_file <- "LakeEnsemblR_10depth.yaml"

# model <- c("FLake", "GLM", "GOTM", "Simstrat",  "MyLake")

model <- c("FLake","GOTM","Simstrat")

export_config(config_file = config_file, model = model)

run_ensemble(config_file = config_file, model = model, parallel = T, return_list = T, ncores = 4)

file.rename(paste0(getwd(),"output"),

            paste0(getwd(),"output_10depth"))

# 10x area ----

config_file <- "LakeEnsemblR_10area.yaml"

# model <- c("FLake", "GLM", "GOTM", "Simstrat",  "MyLake")

model <- c("FLake","GOTM","Simstrat")

export_config(config_file = config_file, model = model)

run_ensemble(config_file = config_file, model = model, parallel = T, return_list = T, ncores = 4)

file.rename(paste0(getwd(),"output"),

            paste0(getwd(),"output_10area"))

# 10x area and 10x depth ----

config_file <- "LakeEnsemblR_10area10depth.yaml"

# model <- c("FLake", "GLM", "GOTM", "Simstrat",  "MyLake")

model <- c("FLake","GOTM","Simstrat")

export_config(config_file = config_file, model = model)

run_ensemble(config_file = config_file, model = model, parallel = T, return_list = T, ncores = 4)

file.rename(paste0(getwd(),"output"),

            paste0(getwd(),"output_10area10depth"))

```
